### 2015-10-22 Article.tex
first commit

%&latex
%%----------------------------------------------------------------------
%% ieeepes_skel.tex
%%
%% Skeleton file for papers for the IEEE Power Engineering Society using
%% package ieeepes.
%%
%% Not copyrighted. Copy this file to a different name and fill in your
%% text.
%%
%% Volker Kuhlmann
%% c/o EEE Dept
%% University of Canterbury
%% Private Bag 4800
%% Christchurch, New Zealand
%% Email: KUHLMAV@ELEC.CANTERBURY.AC.NZ
%%
% 1.3  13Apr99  Updated for ieeepes 4.0.
% 1.2  16Nov95  Fixed discussion, closure. Added summary.
% 1.1  12Nov95  Finished first release.
% 1.02 09Nov95  Option PStimes.
% 1.0  07Nov95  Created.
%%----------------------------------------------------------------------

\documentclass[10pt,twoside%
                ,draft%         % comment this out for final version
        ]{article}

\usepackage[%
        %       psphotos,%      % uncomment those options you want
        %       photofit,%
        %       draft,%
        %       PStimes,%
        ]{ieeepes}
\usepackage{algpseudocode}
\usepackage{algorithm}

\title{Control of Prosumer Networks}

\author{
        N. Gensollen\\
        Institut Mines Telecom\\
        Saclay, France
\and
        V. Gauthier\\
        Institut Mines Telecom\\
        Saclay, France
}

\begin{document}

% This \maketitle command is required from ieeepes version 4.0, to make
% ieeepes work correctly with newer LaTeX versions.
\maketitle

\begin{abstract}
Put the text of your abstract here.
\end{abstract}

\section{Introduction}

Modernizing the power grids and increasing the share of renewables in the production are two substantial objectives of the energetic transition. Because of parallel progresses in communication, data management, and storage, the upcoming emergence of a power grid "2.0", often called smart grid, appears as a cross-disciplinary challenge of the 21st century.

It is believed that these new ways of producing and managing energy, will also impact how end users consume. More precisely, the consumer   of the future is expected to be more involved in the system operation. Using dynamic signals (real time electricity prices for instance), load curves could be shaped to some extend as to agree with the production conditions of the main grid.
Furthermore, the emergence of renewable distributed energy 
 resources (DER) in the distribution networks, possibly hold by individuals,
will completely  
modify the today end user agent model.

These new agents that can both consume and produce are often called prosumers. This decentralized production is believed to lead to more resilient systems that today centralized power grid. Nonetheless, the difficulty of dealing with decentralized production is coupled with the high uncertainty of renewables. Ensure system stability in these conditions appears thus as a complex task.

Relying on the ability of the smart grid to obtain, communicate and analyse data, several solutions could  improve the stability of such complex systems. Among these solutions, demand side management and energy storage are two of the most popular ones. In this paper, we explore the use of storage in a network of prosumers whose power distribution is susceptible to change due to various external causes. More precisely, we are interested in the "best" locations for the storage equipments such that we are able to maintain stability when small perturbations occur. Since all nodes can switch from generator to load and vice versa, the optimal location is susceptible to change with the prosumers power distribution.
Our goal is then to find the one that, on average, provide the best results.

Since charging and discharging a battery is costly, we are particularly interested in the energy necessary to control the system when perturbations occur. The notion of performance of a given set of locations for using storage would then be inversely proportional to the energy used. One problem with this definition is that it depends explicitly on the state of the system and the prosumers power distribution. Besides, as the system size increases, the number of possible location sets increases exponentially. This forbid any algorithms using complete enumeration and evaluation of the sets.
Moreover, we consider  that the electrical lines have maximum capacities, and storage equipments have both a maximum charge / discharge rate and  bounded capacities. Obviously, only solutions satisfying these constraints will be considered as feasible.

 In this paper, we model the prosumer network as a coupled oscillator network. Inspired by [], we use a second order Kuramoto model that can be seen as a simplified version of the underlying power grid dynamics (see Section II). We then use control theory in order to model the impact of the storage equipments on the network (see Section III). Using the relationships between submodular set functions
and the Gramian matrix [], we use a modified version of a greedy algorithm in order to find the smallest locations set that minimizes the average control energy while satisfying all the constraints introduced above (see section IV).
Finally, we show some results in Section V.\section{The Model}

In this section, we introduce the coupled oscillators network model used to simplify the power grid dynamic. The first part is largely inspired by [], and reproduced here for convenience of the reader.

 The objective is to achieve synchronization of the grid at the main frequency $ \Omega = 50 Hz $. Each oscillator i has a phase angle $ \delta_i $ and a frequency $ \dot{\delta}_i $. Therefore, we seek an equilibrium of the form : $ \forall i,\ \dot{\delta}_i = \Omega $. For convenience, we express the dynamic of the oscillators in terms of the deviations from the main frequency : $ \delta_i(t) = \Omega t + \theta_i(t) $. Let $ \omega_i = \dot{\theta_i} $, such that $ \dot{\delta_i} = \Omega + \omega_i $. The equilibrium, in terms of the deviations dynamics, is : $ \forall i,\ \omega_i = 0 $

The next step consists in translating the dynamics of the generators and machines into equations involving the phase angles $ \theta_i $ and the frequencies $ \omega_i$. Generators and machines are composed of a turbine that dissipates energy at a rate proportional to the square of the angular velocity ($P_{diss} = FV \sim V^2 $ ) :

\begin{equation}
  P_{diss, i}(t) = K_{Di}(\dot{\delta_i}(t))^2 
\end{equation}

where $ K_{Di} $ is the dissipation constant of entity i.

Furthermore, it also accumulates kinetic energy at a rate :

\begin{equation} 
P_{acc,i}(t) = \frac{1}{2}I_i\frac{d}{dt} \left( \dot{ \delta_i }(t)^2 \right)
\end{equation}

where $ I_{i}$ is the moment of inertia of entity i. For simplicity, we consider that all entities have the same dissipation constants($K_D$) and moment of inertia (I).

The condition for the power transmission between i and j is that the two devices do not operate in phase : The phase difference between i and j is : $ \delta_j(t) - \delta_i(t) = \Omega t + \theta_j(t) - \Omega t - \theta_i(t) = \theta_j(t) - \theta_i(t) $.

The transmitted power along the line can thus be written as :

\begin{equation}
 P_{transmitted} = - P_{ij}^{MAX} sin( \theta_j - \theta_i )
\end{equation}

with $ P_{ij}^{MAX} $ being the maximum capacity of the line $(i,j)$. If i is connected to more than one other entities, this equation becomes ($\mathcal{N}_i$ being the neighborhood of  i) :

\begin{equation}
  P_{transmitted} = - \sum_{j \in \mathcal{N}_i} P_{ij}^{MAX} sin( \theta_j - \theta_i ) 
\end{equation}

Each entity i is then described by a power balance equation of the type :

\begin{equation}
 P_{S,i}  =  P_{diss,i} + P_{acc,i} + P_{transmitted,i} 
\end{equation}

By substituting and re-arranging the terms, we obtain the following non-linear coupled system of equations :

\begin{equation}
\begin{array}{lll}
P_{S, i}  =  I \Omega \ddot{ \theta_i }  + \left[ I \ddot{ \theta_i } + 2 K_D \Omega \right] \dot{ \theta_i } + K_D \Omega^2 \\+ K_D \dot{ \theta_i }^2 - \sum_{j \in \mathcal{N}_i} P_{ij}^{MAX} sin \left[ \theta_j - \theta_i \right] \end{array}
\end{equation}

These non linear equations are obviously not very practical. We now use some approximations based on the fact that we consider only small deviations from the main frequency : $ \dot{ \delta}_i \sim \Omega $ which means that $ \omega_i = \dot{\theta_i} << \Omega $, such that the squared term $ K_D \dot{\theta}_i^2 $ can be neglected.
Moreover, we assume that the rate at which energy is stored in the kinetic term is much less of the rate at which energy is dissipated by friction, a common condition for mechanical systems : $ \ddot{ \theta }_i  << \dfrac{2 K_D}{I} $

The equation can thus be simplified to :

\begin{equation}
 \ddot{ \theta_i } \sim \psi_i - \alpha \dot{ \theta_i } + \sum_{j\neq i} K_{ij} g_{ji} sin \left[ \theta_j - \theta_i \right] 
\end{equation}

Where $ \alpha = \dfrac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P^i_{source}}{I \Omega} - \frac{K_D \Omega}{I} \right] $, and $ g_{ij} $ is the coefficient of the adjacency matrix G.

The dynamic is still non linear because of the sine coupling. Therefore, we also assume that the phase angle differences are small such that $ sin \left[ \theta_j - \theta_i \right] \sim \theta_j - \theta_i $. By using vector notations, the dynamic can be written in the following form :

\begin{equation}
\ddot{\theta} = \Psi - \alpha \dot{\theta} - KL\theta }
\end{equation}

Where L is the Laplacian matrix of the underlying topology ($ k_i $ is the degree of node i):

\begin{equation}
L_{ij} = \left\{ \begin{array}{lll} k_i\ \ \ \ \,if\ i=j \\ -g_{ij}\  if\ i \neq j \end{array} \right. 
\end{equation}

This is a continuous time second order linear system of $ N $ equations. We first transform this into a continuous time first order linear system of $ 2 N $ equations by introducing the following vector : $ X = \left( \begin{array}{c} \theta \\ \dot{\theta} \end{array} \right)$ :

\begin{equation}
 \dot{X} = \left( \begin{array}{cc} 0 & I \\ - KL & -\alpha I \end{array} \right) X + \left( \begin{array}{c} 0 \\ \Psi \end{array} \right)
\end{equation}

Which can be written in discrete time:

\begin{equation} 
X(t+\Delta t) = M  X(t) +  \left( \begin{array}{c} 0 \\ \Psi \Delta t \end{array} \right)
\end{equation}

With $ M = \left( \begin{array}{cc} I & I \Delta t \\ -KL \Delta t & (1-\alpha \Delta t) I \end{array} \right) $.

By setting $ Y(t) =  \left( \begin{array}{c} X(t) \\ 1 \end{array} \right) $, let the dynamic be :

\begin{equation}
 Y(t+\Delta t) = A Y(t) }
\end{equation}

With transition matrix A :

\begin{equation}
 A =   \left( \begin{array}{ccc} I & I \Delta t & 0 \\ -KL \Delta t & (1-\alpha \Delta t)I & \Psi \Delta t \\ 0&0&1 \end{array} \right)
\end{equation}

This is a linear discrete time system of $ 2 N + 1 $ equations. Note that the transition matrix A encodes all the system parameters, topology, and power distribution.

\section{Control and Submodularity}

\subsection{Submodularity}

In this section we provide a very brief introduction to what submodular functions are, and how their maximization could be achieved in reasonable time.

 A set function $ F:2^{V} \longrightarrow \Re $ defined over a finite set V is said to be submodular if it satisfies :

\begin{equation}
\forall S,T \in V,\ F(S) + F(T) \geq F(S \cup T) + F(S \cap T)
\end{equation}

Another equivalent definition is that for all sets $ X, Y \in V$, such that $ X \subseteq Y$ and for all element $ x \in V \setminus Y$,

\begin{equation}
F(X \cup \{ x \} ) - F(X) \geq F(Y \cup \{ x \} ) - F(Y)
\end{equation}

This basically means that submodular functions exhibit a diminishing return property which makes them particularly interesting for optimization.

Finding the set $ S_k $ of size k that maximizes a set function F is a difficult problem because the number of sets grows exponentially with the number of nodes. Therefore complete enumeration and evaluation is only a feasible solution on very small examples. Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k).

Since the evaluation of F can be costly, a well-known lazy-greedy variation has been proposed by Minoux in 19XX. This smart implementation uses the submodular structure of the marginal gains in order to reduce drastically the number of calls to F. This can result in speedups of several order of magnitudes.

\subsection{Control}

In this section, we introduce basic tools from control theory and relate them to the submodular functions introduced above. Recall that our objective is to maintain a given state in a dynamical system even if perturbations of the power distribution occur.

More precisely, we assume the control of a subset of the nodes that we call driver nodes. In our particular power grid situation, a driver node will be used to model a battery. This control assumption modifies the dynamic in the following way :

\begin{equation}
 Y(t+\Delta t) = A Y(t) + B u(t) 
\end{equation}

Where the control matrix B is an indicator of the driver nodes, and $ u(t) $ represents the control vector input at time t. The system is said to be controllable in T steps is it can be steered from any initial state $ Y_0 $ to any final state $ Y_f $ through a sequence of control inputs $ u(t) $. If the system is controllable, there might be more than one sequence of control inputs that could realize such a thing. Among all these possibilities, we are interested in the one that requires the minimum amount of energy. Let $ \mathcal{E} = \int_{t_0}^{t_0+T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system.

It can be shown that the control input sequence that minimizes the control energy can be written as :

\begin{equation}
 u^{\star}(t) = B^T(A^T)^{T-t-1}W^{-1}(T)\nu(T)
\end{equation}
where $ \nu(T) = Y_f - A^T Y_0 $ is the difference between the desired final state and the final free state, and $W(T) = \sum_{k=0}^{T} A^k B B^T (A^T)^k $ is called the Gramian matrix of the system. The Gramian is a popular tool in control theory since it is deeply linked to the controllability of the system. It can indeed be shown that the system is controllable if and only if the Gramian is not singular, and that its rank indicates the dimension of the controllable subspace. Besides, the minimum control energy associated with the inputs $ u^{\star}(t)$ can be written as :

\begin{equation}
 \mathcal{E}_{min} = \nu(T)^T W^{-1}(T) \nu(T)
\end{equation}

Not surprisingly, the control energy depends on the initial and final states as well as on the inverse of the Gramian. In our prosumer case, we want to find the best battery placement, whose quality should not depend on the initial and final states. Indeed, these states as well as  the power  distribution $ \Psi $ are susceptible to change, and we would like to perform well, on average, whatever the states and distributions. Therefore, we concentrate on  the Gramian as a way to quantify the average controllability. It can indeed be verified that the Gramian does not depend on $\Psi$.

A key point to design a low consumption control is thus to understand the relationships between the control node set and the Gramian matrix. In [], the authors introduce several Gramian-based metrics that have concrete meanings in terms of the system controllability. For example, the trace of the inverse Gramian ($ Tr[ W^{-1}(T)] $) quantifies the average control energy necessary to move the system around the state space. These metrics measures some notions of the control allowed by a given driver node set. But they can be understood the other way around : what drivers should one select as to optimize a given control-based metric ?

Using the Gramian-based metric $ \mathcal{M} $, let $ F(S) =\mathcal{M}(W_S(T))$, where $ W_S(T) $  is the Gramian matrix based on the driver node set S. [] showed that for some metrics, F is a submodular set function. This means that, for these metrics, we can use the greedy heuristic discussed above in order to maximize F.

\section{Control of the Prosumer Network}

\subsection{Controllability Constraints}

 Considering the dynamic of equation (), we notice that the Gramian based on the transition matrix A is never invertible, meaning that the system should not be controllable. It is easy to see that, indeed, we do not have full control because of the last row of the matrix. Nevertheless, as this row was only added in order to include the power distribution inside A, we do not need to control the system along this "fake" dimension. That is, we only seek the control of the system in the first $ 2N $ dimensions of the space (N phase angles $ \theta_i$ and N frequencies $\omega_i$). Therefore, whenever $ rank[ W(T) ] = 2N$, we will use the Moore-Penrose pseudo inverse of the gramian $ W^\dagger(T) $.

\subsection{Flow Constraints}

Recall that the simplified dynamic (see section II) implies that, the power transmitted on line $ (i,j)$ is $ P_{i \longrightarrow j} = -P_{ij}^{MAX} \left( \theta_j(t) - \theta_i(t) \right) $. Therefore, we write the flow constraints at time t as :

\begin{equation}
\forall (i,j) \in V^2,\ \ g_{ij} \left|\ \theta_j(t) - \theta_i(t)\ \right| \leq 1
\end{equation}
If this constraint is verified for all instant t in the control time range, then the control inputs satisfy the flow constraints.

\subsection{Power Distribution Constraints}
As explained above, we wish to obtain results that do not depend on the power distribution $ \Psi $ of the prosumers. However, in order to achieve synchronization along with feasible power flows, we need to define some constraints on this distribution. As shown in [], a condition for achieving synchronisation in a coupled oscillators network is $ \Vert L^{\dagger}\omega \Vert_{\infty} \leq sin(\gamma) $, where L is the Laplacian matrix of the network, $ \omega $ is the vector of the natural frequencies of the oscillators, and $ \Vert x \Vert_{\infty} = max \left| x_i - x_j \right| $. If this condition is satisfied, the oscillators will synchronize at the common frequency $ \omega_{SYNC} $ with the phase lock $ g_{ij}\left| \theta_i - \theta_j \right| \leq    \gamma \in[0,\frac{\pi}{2}],\ \forall (i,j) \in V^2 $.

Therefore, we can write the following synchronization constraint :

\begin{equation}
\label{synchro_constraint}
\Vert \left( L \circ P^{MAX} \right)^{\dagger} \Psi \Vert_{\infty} \leq 1
\end{equation}

Where $ A \circ B $ represents the Hadamard product between matrices A and B. If constraint \ref{synchro_constraint} is satisfied, the oscillators will synchronize to a common frequency $ \omega_{SYNC} = \frac{\sum_{k=1}^{N}\Psi_k}{\sum_{k=1}^{N}\alpha_k} $, and the flows at steady state will be feasible. Note that $\omega_{SYNC} $ should be zero in order to synchronize to the main frequency $\Omega$ (recall that the phase angles and frequencies represent deviations from $ \Omega$). This gives the following intuitive zero sum constraint :

\begin{equation}
\label{zero_sum_constraint}
\sum_{k=1}^{N} \Psi_k = 0
\end{equation}

Constraint \ref{zero_sum_constraint} simply states that production should match demand in order for the system to remain stable.

\subsection{Battery Constraints}

Let i be a battery. In the simple model we adopt here, it is characterised by its maximum charge/discharge rate $r_i$, the amount of energy stored at time t $ \Lambda_i(t) $, and its maximum capacity $ \Lambda_{i, MAX} $. For simplicity, we will consider all maximum charge/discharge rates to be the same ($r$), and all maximum capacities to be the same $\Lambda_{Max}$. Furthermore, we denote by $ \Lambda(t) $ the vector of energy level at time t.

Since the control input vector $ u^{\star}(t) $ specifies the control inputs, the energy level dynamic of the batteries can be written as :

\begin{equation}
\Lambda(t+\Delta t) = \Lambda(t) - u^{\star}(t) I \Omega 
\end{equation}

Which can also be written as :

\begin{equation}
\Lambda(t)  = \Lambda(0) - I \Omega \sum_{k=0}^{t}u^{\star}(k)
\end{equation}

Where $ \Lambda(0) $ is the vector of initial levels in the batteries.

Obviously, the energy level in a battery is bounded. This leads to the following set of constraints :

\begin{equation}
\label{level_constraints}
 \forall t, \ \ 0 \leq \Lambda(t) \leq \Lambda_{Max}
\end{equation}

Finally, during each time slot, the battery cannot charge or discharge at a rate higher than $ r $ :

\begin{equation}
\label{rate_constraints}
 \forall t,\ \ \left|\ u^{\star}(t)I \Omega\ \right| \leq r 
 \end{equation}

If, for a given $ u^{\star} $, the constraints of eq. \ref{level_constraints} and \ref{rate_constraints} are satisfied, then the control inputs satisfy the storage constraints.

\subsection{Algorithm}

We propose the following greedy algorithm in order to find the smallest set of locations that minimizes the average control energy and satisfy the constraints. We assume here that the have a given initial state $ Y_0 $ and a desired final state $ Y_f$ that we want to reach in $ T $ time steps. The algorithm \ref{algo_1} starts with an empty set and, as long as that constraints are not satisfied, it increases the size of the set. For a given size k, the set $ S_k $ is found using lazy greedy submodular optimization of a Gramian-based set function F. In practice we use $ F(S) = -Tr \left[ W_S^{\dagger}(T) \right] $ which quantifies the average energy required to move the system around the controllable subspace. The algorithm \ref{algo_1} returns then the tuple $ (k,S_k,u^{\star}) $ such that $ S_k $ is the smallest set that minimizes the average control energy and, in this case, satisfies the constraints with control $u^{\star}$.

\begin{algorithm}
\label{algo_1}
        \begin{algorithmic}
                \While {\small $ Not\ \left\{ \begin{array}{lll} rank \left[ W_{S_k}(T) \right] < 2N\ and \\ \forall t,\ \forall (i,j) \in V^2,\ g_{ij} \left|\theta_j(t)-\theta_i(t) \right| \leq 1\ and \\ \forall t,\ 0 \leq \Lambda(t) \leq \Lambda_{MAX}\ and \\ \forall t,\ u^{\star}(t)I\Omega \leq r\end{array} \right. $ }
    \State $k\ \ \ \ \ \ \ \ \ \ \ \gets k+1$
    \State $ S_k \ \ \ \ \ \ \ \ \ \gets Subopt(F, k)$
    \State $ u^{\star}(t)\ \ \ \ \ \ \gets OptControl(S_{k})$ 
    \State $ Y(t), \Lambda(t) \gets Simulation(A,B,T,Y_0,Y_f,u^{\star}) $

        \end{algorithmic}
\end{algorithm}
\normalsize
Since changing $ \Psi $, $Y_0$, or $Y_f$ does not change the Gramian, variations in the power distribution or initial and final states will not change the $S_k$ whatever k. However, they have an impact on the flows and battery levels since they appear explicitly in the control energy.

\section{Results}

\bibliography{FilenameOfYourBibliography}

\end{document}

%%
%% EOF ieeepes_skel.tex
%%----------------------------------------------------------------------

### 2015-10-26 Article.tex
update

@@
 \begin{abstract}
-Put the text of your abstract here.
+The ability to control a given network by  dynamically injecting inputs offers ways to make sure that it stays in stable regions.
+In the case of smart grids, where end users exhibit complex behaviors and renewable production is quite unstable, control of the system  is paramount.
+In this paper, we study the case of a network composed of entities called prosumers.
+These agents have the ability to both consume and produce according to external conditions.
+We first model and simplify the underlying dynamic of such network using a second order coupled oscillators network model.
+Under some conditions, the system synchronize to a common frequency.
+However, in case of perturbations in the power distribution, the system might lose synchrony and require control to bring it back to the stable state.
+In this situation, control can be seen as energy  absorbed or injected in the system at specific locations.
+Moreover, the power outputs of the prosumers are susceptible to change such that loads and generators are not fixed.
+In this context, it is important to select the right subset
+of nodes that yields, on average, the cheapest control in terms of energy.
+
 \end{abstract}
@@
 These new agents that can both consume and produce are often called prosumers.
-This decentralized production is believed to lead to more resilient systems that today centralized power grid.
+This decentralized production is believed to lead to more resilient systems than today centralized power grid.
 Nonetheless, the difficulty of dealing with decentralized production is coupled with the high uncertainty of renewables.
@@
 Among these solutions, demand side management and energy storage are two of the most popular ones.
-In this paper, we explore the use of storage in a network of prosumers whose power distribution is susceptible to change due to various external causes.
-More precisely, we are interested in the "best" locations for the storage equipments such that we are able to maintain stability when small perturbations occur.
-Since all nodes can switch from generator to load and vice versa, the optimal location is susceptible to change with the prosumers power distribution.
+In this paper, we explore the use of storage in a network of prosumers whose power outputs are susceptible to change due to various external causes.
+More precisely, we are interested in the "best" locations for the storage equipments such that we are able to maintain stability when  perturbations occur.
+Since all nodes can switch from generator to load and vice versa, the optimal location is susceptible to change with the prosumers power output.
 Our goal is then to find the one that, on average, provide the best results.
@@
 
-Where $ \alpha = \dfrac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P^i_{source}}{I \Omega} - \frac{K_D \Omega}{I} \right] $, and $ g_{ij} $ is the coefficient of the adjacency matrix G.
+Where $ \alpha = \dfrac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $, and $ g_{ij} $ is the coefficient of the adjacency matrix G.
+By working in a rotating frame, we can simplify $ \psi_i=\frac{P_{S,i}}{I \Omega} $.
 
@@
 \label{synchro_constraint}
-\Vert \left( L \circ P^{MAX} \right)^{\dagger} \Psi \Vert_{\infty} \leq 1
+\Vert \left( L \circ P^{MAX} \right)^{\dagger} \Psi \Vert_{\infty} \leq sin(1)
 \end{equation}
@@
 Where $ A \circ B $ represents the Hadamard product between matrices A and B.
-If constraint \ref{synchro_constraint} is satisfied, the oscillators will synchronize to a common frequency $ \omega_{SYNC} = \frac{\sum_{k=1}^{N}\Psi_k}{\sum_{k=1}^{N}\alpha_k} $, and the flows at steady state will be feasible.
+If constraint \ref{synchro_constraint} is satisfied, the oscillators will synchronize to a common frequency $ \omega_{SYNC} = \frac{\sum_{k=1}^{N}\Psi_k}{\sum_{k=1}^{N}\alpha_k} = \frac{\sum_{k=1}^{N}P_{S,k}}{I\Omega N \alpha}$, and the flows at steady state will be feasible.
 Note that $\omega_{SYNC} $ should be zero in order to synchronize to the main frequency $\Omega$ (recall that the phase angles and frequencies represent deviations from $ \Omega$).
@@
 \label{zero_sum_constraint}
-\sum_{k=1}^{N} \Psi_k = 0
+\sum_{k=1}^{N} P_{S,k} = 0
 \end{equation}
@@
 \begin{equation}
+ \forall t, \ \ 0 \leq \Lambda(t) \leq \Lambda_{MAX}
+\end{equation}
+
+Which can be written in terms of the control inputs :
+
+\begin{equation}
 \label{level_constraints}
- \forall t, \ \ 0 \leq \Lambda(t) \leq \Lambda_{Max}
+\forall t,-\frac{\Lambda(0)}{I \Omega} \leq -\sum_{k=0}^{t}u^{\star}(k) \leq \frac{ \Lambda_{MAX} - \Lambda(0)}{I \Omega}
 \end{equation}
@@
 We propose the following greedy algorithm in order to find the smallest set of locations that minimizes the average control energy and satisfy the constraints.
-We assume here that the have a given initial state $ Y_0 $ and a desired final state $ Y_f$ that we want to reach in $ T $ time steps.
+We assume here that the system is in a given initial state $ Y_0 $ and we want to reach a desired final state $ Y_f$  in $ T $ time steps.
 The algorithm \ref{algo_1} starts with an empty set and, as long as that constraints are not satisfied, it increases the size of the set.
@@
         \begin{algorithmic}
-                \While {\small $ Not\ \left\{ \begin{array}{lll} rank \left[ W_{S_k}(T) \right] < 2N\ and \\ \forall t,\ \forall (i,j) \in V^2,\ g_{ij} \left|\theta_j(t)-\theta_i(t) \right| \leq 1\ and \\ \forall t,\ 0 \leq \Lambda(t) \leq \Lambda_{MAX}\ and \\ \forall t,\ u^{\star}(t)I\Omega \leq r\end{array} \right.
+                \While {\small $ Not\ \left\{ \begin{array}{lll} rank \left[ W_{S_k}(T) \right] < 2N\ and \\ \forall t,\forall (i,j) \in V^2, g_{ij} \left| \theta_j(t) - \theta_i(t) \right| \leq 1\  and \\ \forall t,-\frac{\Lambda(0)}{I \Omega} \leq -\sum_{k=0}^{t}u^{\star}(k) \leq \frac{ \Lambda_{MAX} - \Lambda(0)}{I \Omega}and \\ \forall t,\ \left| u^{\star}(t) \right| \leq \frac{r}{I\Omega}\end{array} \right.
 $ }
@@
     \State $ u^{\star}(t)\ \ \ \ \ \ \gets OptControl(S_{k})$
-    \State $ Y(t), \Lambda(t) \gets Simulation(A,B,T,Y_0,Y_f,u^{\star}) $
 
@@
 \normalsize
-Since changing $ \Psi $, $Y_0$, or $Y_f$ does not change the Gramian, variations in the power distribution or initial and final states will not change the $S_k$ whatever k.
-However, they have an impact on the flows and battery levels since they appear explicitly in the control energy.
 
 \section{Results}
+In this section, we provide some results in order to illustrate the work described above.
+We use a small 10 prosumers example, connected with a random topology.
+The line maximum capacities are selected  according to a normal distribution $\mathcal{N}(\mu_l,\sigma_l)$.
+The power distribution is chosen as a zero mean normal distribution $ \mathcal{N}(0,\sigma_p) $, and the samples should satisfy the constraints \ref{synchro_constraint} and \ref{zero_sum_constraint}.
+Obviously, $\mu_l$, $\sigma_l$, and $\sigma_p$ should be chosen such that, without any perturbation, the lines are not already overloaded.
+
+Once the synchronized state is reached, we apply a perturbation to the power distribution such that there is a mismatch between production and demand (constraint \ref{zero_sum_constraint} is not satisfied).
+We denote by $d$ the delay between the perturbation and the start of the control phase.
+$d$ depends on how fast the perturbation can be identified and located, the control inputs calculated and communicated without error through the communication network.
+
+ If the driver set is fixed, we expect that the higher $ d$, the higher the control energy required to drive the system back to the synchronized state.
+Eventually, if $d$ is too large, the minimum energy control inputs will result in some constraints being broken (battery levels, line flows, and so on...).
+
+Figure X illustrates the evolution of the state variables (phase angles and frequencies).
+As visible, when d increases, the system state deviates more from the equilibrium and results in more energy being used to steer it back to synchrony.
+Figure XX shows how the global control energy evolves with the delay $ d$.
+The red dot indicates the maximum delay tolerable by the system before a constraint is broken.
+Here, some batteries become empty as shown in figure XXX.
+
+We next investigate how the driver node set selected with algorithm X performs against other driver sets of the same size.
+Since, the example is relatively small, we can enumerate all these sets, and filter out the ones that do not meet all the constraints.
+We then sample the power distribution of the prosumers and compare the resulting mean  control energies for all the sets.
 

### 2015-10-28 Article.tex
update

@@
-%&latex
-%%----------------------------------------------------------------------
-%% ieeepes_skel.tex
-%%
-%% Skeleton file for papers for the IEEE Power Engineering Society using
-%% package ieeepes.
-%%
-%% Not copyrighted. Copy this file to a different name and fill in your
-%% text.
-%%
-%% Volker Kuhlmann
-%% c/o EEE Dept
-%% University of Canterbury
-%% Private Bag 4800
-%% Christchurch, New Zealand
-%% Email: KUHLMAV@ELEC.CANTERBURY.AC.NZ
-%%
-% 1.3  13Apr99  Updated for ieeepes 4.0.
-% 1.2  16Nov95  Fixed discussion, closure. Added summary.
-% 1.1  12Nov95  Finished first release.
-% 1.02 09Nov95  Option PStimes.
-% 1.0  07Nov95  Created.
-%%----------------------------------------------------------------------
+\documentclass[conference]{IEEEtran}
 
-\documentclass[10pt,twoside%
-                ,draft%         % comment this out for final version
-        ]{article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Package
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{ifpdf}
+\ifCLASSINFOpdf
+  \usepackage[pdftex]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../pdf/}{../jpeg/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\else
+  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
+  % will default to the driver specified in the system graphics.cfg if no
+  % driver is specified.
+  \usepackage[dvips]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../eps/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  \DeclareGraphicsExtensions{.eps}
+\fi
 
-\usepackage[%
-        %       psphotos,%      % uncomment those options you want
-        %       photofit,%
-        %       draft,%
-        %       PStimes,%
-        ]{ieeepes}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage[french, english]{babel}
+\usepackage{cite}
+\usepackage{caption}
+\usepackage{amsmath,amsfonts,amssymb}
+\usepackage{subcaption}
+\usepackage{array}
+\usepackage{float}
+\usepackage{multicol}
 \usepackage{algpseudocode}
@@
 
+\begin{document}
+\pagestyle{plain}
 \title{Control of Prosumer Networks}
 
-\author{
-        N. Gensollen\\
-        Institut Mines Telecom\\
-        Saclay, France
-\and
-        V. Gauthier\\
-        Institut Mines Telecom\\
-        Saclay, France
-}
+\author{Nicolas Gensollen, Monique Becker, Vincent Gauthier and Michel Marot  \\
+\IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
+Telecom SudParis/Institut Mines Telecom\\
+Email: \{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu}}
 
-\begin{document}
-
-% This \maketitle command is required from ieeepes version 4.0, to make
-% ieeepes work correctly with newer LaTeX versions.
 \maketitle
@@
 Moreover, the power outputs of the prosumers are susceptible to change such that loads and generators are not fixed.
-In this context, it is important to select the right subset
-of nodes that yields, on average, the cheapest control in terms of energy.
-
+In this context, it is important to select the right subset of nodes that yields, on average, the cheapest control in terms of energy.
 \end{abstract}
 
+\IEEEpeerreviewmaketitle
+
 \section{Introduction}
+\label{sec:introduction}
 
 Modernizing the power grids and increasing the share of renewables in the production are two substantial objectives of the energetic transition.
-Because of parallel progresses in communication, data management, and storage, the upcoming emergence of a power grid "2.0", often called smart grid, appears as a cross-disciplinary challenge of the 21st century.
+Because of parallel progresses in communication, data management, and storage, the upcoming emergence of a power grid "2.0", often called smart grid, appears as a cross-disciplinary challenge of the 21st century \cite{Ramchurn}.
 
@@
 Using dynamic signals (real time electricity prices for instance), load curves could be shaped to some extend as to agree with the production conditions of the main grid.
-Furthermore, the emergence of renewable distributed energy 
- resources (DER) in the distribution networks, possibly hold by individuals,
-will completely  
-modify the today end user agent model.
+Furthermore, the emergence of renewable distributed energy resources (DER) in the distribution networks, possibly hold by individuals,
+will completely modify the today end user agent model.
 
-These new agents that can both consume and produce are often called prosumers.
+These new agents that can both consume and produce are often called prosumers \cite{Rathnayaka2012}.
 This decentralized production is believed to lead to more resilient systems than today centralized power grid.
@@
  In this paper, we model the prosumer network as a coupled oscillator network.
-Inspired by [], we use a second order Kuramoto model that can be seen as a simplified version of the underlying power grid dynamics (see Section II).
-We then use control theory in order to model the impact of the storage equipments on the network (see Section III).
-Using the relationships between submodular set functions
-and the Gramian matrix [], we use a modified version of a greedy algorithm in order to find the smallest locations set that minimizes the average control energy while satisfying all the constraints introduced above (see section IV).
-Finally, we show some results in Section V.\section{The Model}
+Inspired by \cite{Filatrella2008}, we use a second order Kuramoto model that can be seen as a simplified version of the underlying power grid dynamics (see Section \ref{sec:The_Model}).
+We then use control theory \cite{Liu2015} in order to model the impact of the storage equipments on the network (see Section \ref{sec:control_and_submodularity}).
+Using the relationships between submodular set functions and the Gramian matrix \cite{Summers2014}, we use a modified version of a greedy algorithm \cite{Minoux} in order to find the smallest locations set that minimizes the average control energy while satisfying all the constraints (see section \ref{sec:Control_prosumer}).
+Finally, we show some results in Section \ref{sec:Results}.
+
+\section{The Model}
+\label{sec:The_Model}
 
 In this section, we introduce the coupled oscillators network model used to simplify the power grid dynamic.
-The first part is largely inspired by [], and reproduced here for convenience of the reader.
+The detailed derivation of the second order dynamics can be found in \cite{Filatrella2008}, and the key points are reproduced here for the convenience of the reader.
 
@@
 We now use some approximations based on the fact that we consider only small deviations from the main frequency : $ \dot{ \delta}_i \sim \Omega $ which means that $ \omega_i = \dot{\theta_i} << \Omega $, such that the squared term $ K_D \dot{\theta}_i^2 $ can be neglected.
-Moreover, we assume that the rate at which energy is stored in the kinetic term is much less of the rate at which energy is dissipated by friction, a common condition for mechanical systems : $ \ddot{ \theta }_i  << \dfrac{2 K_D}{I} $
+Moreover, we assume that the rate at which energy is stored in the kinetic term is much less of the rate at which energy is dissipated by friction, a common condition for mechanical systems : $ \ddot{ \theta }_i  << \frac{2 K_D}{I} $
 
@@
 
-Where $ \alpha = \dfrac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $, and $ g_{ij} $ is the coefficient of the adjacency matrix G.
+Where $ \alpha = \frac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $, and $ g_{ij} $ is the coefficient of the adjacency matrix G.
 By working in a rotating frame, we can simplify $ \psi_i=\frac{P_{S,i}}{I \Omega} $.
@@
 \begin{equation}
-\ddot{\theta} = \Psi - \alpha \dot{\theta} - KL\theta }
+\ddot{\theta} = \Psi - \alpha \dot{\theta} - KL\theta
 \end{equation}
@@
 \begin{equation}
- Y(t+\Delta t) = A Y(t) }
+ Y(t+\Delta t) = A Y(t)
 \end{equation}
@@
 \section{Control and Submodularity}
+\label{sec:control_and_submodularity}
 
@@
 In this section we provide a very brief introduction to what submodular functions are, and how their maximization could be achieved in reasonable time.
+More information can be found in \cite{Krause2014}.
 
@@
 Therefore complete enumeration and evaluation is only a feasible solution on very small examples.
-Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k).
+Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k) \cite{Krause2014}.
 
-Since the evaluation of F can be costly, a well-known lazy-greedy variation has been proposed by Minoux in 19XX.
-This smart implementation uses the submodular structure of the marginal gains in order to reduce drastically the number of calls to F.
+Since the evaluation of F can be costly, a well-known lazy-greedy variation has been proposed by \cite{Minoux}.
+This smart implementation uses the submodular structure of the marginal gains in order to reduce the number of calls to F.
 This can result in speedups of several order of magnitudes.
@@
 In this section, we introduce basic tools from control theory and relate them to the submodular functions introduced above.
-Recall that our objective is to maintain a given state in a dynamical system even if perturbations of the power distribution occur.
 
-More precisely, we assume the control of a subset of the nodes that we call driver nodes.
-In our particular power grid situation, a driver node will be used to model a battery.
-This control assumption modifies the dynamic in the following way :
+We assume the control of a subset of the nodes that we call driver nodes.
+In our particular power grid situation, a driver node can be seen as a prosumer where some additional equipment (battery, backup generator,...) will be used as to inject or absorb energy from the system.
+This control assumption modifies the system dynamic :
 
@@
 Where the control matrix B is an indicator of the driver nodes, and $ u(t) $ represents the control vector input at time t.
-The system is said to be controllable in T steps is it can be steered from any initial state $ Y_0 $ to any final state $ Y_f $ through a sequence of control inputs $ u(t) $.
-If the system is controllable, there might be more than one sequence of control inputs that could realize such a thing.
-Among all these possibilities, we are interested in the one that requires the minimum amount of energy.
+The system is said to be controllable in T steps if it can be steered from any initial state $ Y_0 $ to any final state $ Y_f $ through a sequence of control inputs $ u(t) $.
+If the system is controllable, there might be more than one sequence of control inputs that could achieve such a thing.
+Among all these possibilities, we are interested in the one that requires the minimum amount of energy \cite{Yan2012}.
 Let $ \mathcal{E} = \int_{t_0}^{t_0+T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system.
 
-It can be shown that the control input sequence that minimizes the control energy can be written as :
+It can be shown \cite{Liu2015} that the control input sequence that minimizes the control energy can be written as :
 
@@
 \end{equation}
+
 where $ \nu(T) = Y_f - A^T Y_0 $ is the difference between the desired final state and the final free state, and $W(T) = \sum_{k=0}^{T} A^k B B^T (A^T)^k $ is called the Gramian matrix of the system.
@@
 \section{Control of the Prosumer Network}
+\label{sec:Control_prosumer}
 
@@
 	\label{algo_1}
+
         \begin{algorithmic}
-                \While {\small $ Not\ \left\{ \begin{array}{lll} rank \left[ W_{S_k}(T) \right] < 2N\ and \\ \forall t,\forall (i,j) \in V^2, g_{ij} \left| \theta_j(t) - \theta_i(t) \right| \leq 1\  and \\ \forall t,-\frac{\Lambda(0)}{I \Omega} \leq -\sum_{k=0}^{t}u^{\star}(k) \leq \frac{ \Lambda_{MAX} - \Lambda(0)}{I \Omega}and \\ \forall t,\ \left| u^{\star}(t) \right| \leq \frac{r}{I\Omega}\end{array} \right.
-$ }
+                \While{  \small $ Not\ \left\{ \begin{array}{lll} 
+                                              rank \left[ W_{S_k}(T) \right] < 2N\ and \\ 
+                                              \forall t,\forall (i,j) \in V^2, g_{ij} \left| \theta_j(t) - \theta_i(t) \right| \leq 1\  and \\
+                                              \forall t,-\frac{\Lambda(0)}{I \Omega} \leq -\sum_{k=0}^{t}u^{\star}(k) \leq \frac{ \Lambda_{MAX} - \Lambda(0)}{I \Omega}and \\ 
+                                              \forall t,\ \left| u^{\star}(t) \right| \leq \frac{r}{I\Omega}
+
+                                              \end{array} 
+                                              \right. $ }
     \State $k\ \ \ \ \ \ \ \ \ \ \ \gets k+1$
     \State $ S_k \ \ \ \ \ \ \ \ \ \gets Subopt(F, k)$
-    \State $ u^{\star}(t)\ \ \ \ \ \ \gets OptControl(S_{k})$
-
+    \State $ u^{\star}(t)\ \ \ \ \ \ \gets OptControl( S_{k} )$ 
+\EndWhile
         \end{algorithmic}
@@
 \section{Results}
+\label{sec:Results}
+
 In this section, we provide some results in order to illustrate the work described above.
@@
 
+\begin{figure}
+\label{fig:frequecies}
+\includegraphics[scale=.38]{frequencies}
+\caption{Time series of the N frequencies $\omega_i$ (phase angles are not shown for clarity).
+$d$ is the delay between the perturbation and the begining of the control phase that bring the system bak in the synchronized state in T time steps.
+Here, $d$ was chosen quite large for the clarity of the figure.
+Note that, if the perturbation is still active at this point, the system still needs control in order to stay synchronized (batteries should still compensate the power imbalance)}
+\end{figure}
+
 Once the synchronized state is reached, we apply a perturbation to the power distribution such that there is a mismatch between production and demand (constraint \ref{zero_sum_constraint} is not satisfied).
-We denote by $d$ the delay between the perturbation and the start of the control phase.
+We denote by $d$ the delay between the perturbation and the start of the control phase (see figure \ref{fig:frequecies}).
 $d$ depends on how fast the perturbation can be identified and located, the control inputs calculated and communicated without error through the communication network.
@@
 
-Figure X illustrates the evolution of the state variables (phase angles and frequencies).
-As visible, when d increases, the system state deviates more from the equilibrium and results in more energy being used to steer it back to synchrony.
-Figure XX shows how the global control energy evolves with the delay $ d$.
-The red dot indicates the maximum delay tolerable by the system before a constraint is broken.
-Here, some batteries become empty as shown in figure XXX.
+\begin{figure}
+\label{fig:energy_vs_delay}
+\includegraphics[scale=.38]{energy_vs_delay2}
+\caption{ Control energy used in order to bring the system back to the synchronized state when the delay between perturbation and control increases, while the driver node set is fixed.
+The blue dotted curve shows the results for $ S_k^{\star} $ our algorithm \ref{algo_1}.
+The red line shows (average and standard deviation) the results when the driver set is sampled uniformly from all possible sets of the size $ k $.}
+\end{figure}
 
-We next investigate how the driver node set selected with algorithm X performs against other driver sets of the same size.
-Since, the example is relatively small, we can enumerate all these sets, and filter out the ones that do not meet all the constraints.
-We then sample the power distribution of the prosumers and compare the resulting mean  control energies for all the sets.
+As visible on figure \ref{fig:energy_vs_delay}, when d increases, the system deviates more from the equilibrium and this results in more energy being used to steer it back to synchrony.
+We also investigate here how the driver node set selected with algorithm \ref{algo_1} performs against other driver sets of the same size.
+we use a Monte-Carlo sampling of the drivers sets of the same size as our solution, and filter out the ones that do not meet all the constraints.
+Figure \ref{fig:energy_vs_delay} shows that our solution (blue dotted curve) uses less energy than other driver sets of the same size (red curve).
 
-\bibliography{FilenameOfYourBibliography}
+\section{Conclusion}
+\label{sec:Conclusion}
+C'est fini !
+
+\bibliographystyle{IEEEtran}  
+\bibliography{Article}
 
 \end{document}
-
-%%
-%% EOF ieeepes_skel.tex
-%%----------------------------------------------------------------------

### 2016-02-09 Article.tex
big update

@@
 
-In this section, we provide some results in order to illustrate the work described above.
-We use a small 10 prosumers example, connected with a random topology.
-The line maximum capacities are selected  according to a normal distribution $\mathcal{N}(\mu_l,\sigma_l)$.
-The power distribution is chosen as a zero mean normal distribution $ \mathcal{N}(0,\sigma_p) $, and the samples should satisfy the constraints \ref{synchro_constraint} and \ref{zero_sum_constraint}.
-Obviously, $\mu_l$, $\sigma_l$, and $\sigma_p$ should be chosen such that, without any perturbation, the lines are not already overloaded.
+For a direct illustrative application, we chose a small system of 15 nodes connected with an erdos-renyi topology.
+The power distribution and the capacities of the lines are selected randomly such that the system is at equilibrium.
+Therefore, the production matches the consumption, no line is overloaded and all nodes are synchronized to the main frequency.
+We use the greedy algorithm (X) based on the trace of the inverse Gramian as to select the driver set $S$.
+At a given time $ t_k $, we introduce a small perturbation in the form of a mismatch between the production and consumption : $P_i(t_k) = P_i(t_{k-1}) + \delta P$ for some random node i.
+Consequently, the frequencies $ \dot{ \theta }_i \forall i $ of the oscillators start to deviate from the synchronized state.
+At time $ t_k + d $, we apply to the drivers the optimal control of equation (X) such that the system is brought to the synchronized state at time $ t_k + d + T $, where T is the control time.
+At this time, if the mismatch between production and consumption is still active, we maintain synchrony by using optimal control with $T=1$.
+Figure 1 shows the evolution of a subset (for clarity) of the nodes' frequencies during the control phase.
 
@@
 \end{figure}
+
+The purpose of selecting the drivers according to a Gramian related metric is to minimize, on average, the amount of control energy needed.
+In other words, if we select the drivers randomly, we expect to need, on average, more energy to control the system.
+In figure 2, we draw 200 example systems of 15 nodes with random topologies, power distributions, and line capacities.
+These systems are initially at equilibrium and are perturbed as explained previously.
+For each system we find the driver set S using algorithm (X) and we draw a random driver set $S'$ such that $S$ and $S'$ have the same size, and such that $S'$ also provide full control of the system.
+We then compute the control signals and the corresponding control energies $\mathcal{E}_S$ and $\mathcal{E}_{S'}$ for both $S$ and $S'$.
+Each system can therefore be represented as a point with coordinates $(|S|, \mathcal{E}_S)$.
+On figure 2, we plot the densities of these points for the random case (left pannel) and the optimized case (right pannel).
+Regions in dark red contains a lot of points while regions in dark blue have very few points.
+Figure 2 shows that most of the systems considered tend to be controlled by the drivers of algorithm X with less energy and/or drivers than for the random case.
+
+\begin{figure}
+\label{fig:densities}
+\includegraphics[scale=.3]{rr}
+\caption{test}
+\end{figure}
+
+We investigate next how the topology of the grid affects the size of the control set.
+We consider first the case of an erdos renyi topology with probability of connection p.
+We show on Figure 3 how the size of the driver set evolves with p for different Gramian based metrics and for two levels of constraints :
+
+\begin{itemize}
+\item Level 1 : full control
+\item Level 2 : full control + line capacities constraints + battery constraints (see algorithm X)
+\end{itemize}
+
+\begin{figure}
+\label{fig:erdos_renyi}
+\includegraphics[scale=.3]{graphique_1}
+\caption{ $ n_D $ against link probability p for erdos-renyi topologies ($ N=100 $).
+Curves are averaged over 100 realizations.
+The top three curves show the results for the three metrics taken into consideration when all constraints are considered (see Level 2 in the main text).
+The bottom three curves exhibit the results for the same metrics when only the full controllability constraint is considered (see Level 1 in the main text).}
+\end{figure}
+
+When p is close to zero, nodes tend to be very poorly connected such that we need to control almost all nodes in the grid.
+As p increases, the connectivity of the graph augments and the number of drivers decreases.
+At some point, the connectivity of the graph starts to harm its controllability, and more drivers are needed.
+This effect is in accordance with the litterature although it is less marked here because the topology has less influence in our model (recall that matrix A in the dynamics is not the adjacency matrix of the graph but a block matrix where one of the block is the Laplacian matrix of the graph).
+As expected, the driver sets for the level 2 of constraints are larger than for level 1 (for all metrics) because we impose far more constraints.
+
+Because transmission power grids are spatially and politically constrained (each contry has its own transmission grid which is interconnected with neighboring grids at some specific locations), we study the control in clustered graphs.
+More precisely, we use a block model $(N, N_{clusters}, p_{in}, p_{out} ) $ to generate random topologies where $ N $ is the number of nodes, $N_{clusters} $ is the number of clusters, $p_{in} $ is the probability that two nodes within the same cluster are linked, and $ p_{out} $ is the probability that two nodes in two distinct clusters are connected.
+
+\begin{figure}
+\label{fig:block_model_1}
+\includegraphics[scale=.25]{block_model_3}
+\caption{$n_D$ against $N_{clusters} $ for different values of $ p_{in} $ and $ p_{out} = 0.1 $. $ N=200 $ and curves are averaged over 100 realizations.}
+\end{figure}
+
+Figure 4 shows how the number of drivers evolves with the number of clusters for different values of $ p_{in} $ when $ p_{out} $ is fixed ($p_{out} =0.1$).
+For small values of $ p_{in} \sim p_{out} $ clusters are poorly marked, and the connectivity is low.
+The number of drivers in these conditions is large and increases slowly when the number of clusters augments.
+As $ p_{in} $ increases, the clusters becomes more densily connected such that within a cluster less drivers are required in order to control it.
+As the number of clusters grows, more drivers are required.
+For large values of $p_{in} $, the behavious is more complex.
+We see indeed that for $p_{in} = 0.9 $ and $N_{clusters} = 2$, almost $ 94\% $ of the nodes are needed for control, but this quantity first decreases with the addition of a few more clusters ($ 71\% $ for $ N_{clusters}=5 $), before increasing.
+This behavior can be explained by the fact that controlling a very dense network requires, in our settings, a very large portion of nodes to be drivers.
+Controlling one cluster alone thus requires to control almost all nodes within this cluster.
+Nevertheless, when a relatively small (compared to the number of nodes in the graph) number of clusters are interconnected with a few links, nodes in one cluster are able to control nodes in other clusters to which they are connected.
+As the number of clusters grows, the global connectivity increases rapidely, such that the number of drivers also rises.
+We show this behavior in more details in figure 5, where the number of drivers is plotted against $ p_{in} $ for $N=200$ and $N_{clusters} \in \{2,4,5\} $.
+For $N_{clusters} = 2 $ and $ p_{in} = 0.1 $ the number of drivers is large since the graph is poorly connected.
+When $p_{in} $ increases, $n_D$ decreases until a minimum value $n_D \sim 0.64 $ at $ p_{in} \sim 0.6 $.
+After this point, $n_D$ augments with $p_{in}$.
+For $N_{clusters} = 5 $, we do not see this behavior : $ n_D $ decreases continuously as $p_{in} $ increases as expected from the curves of figure 4.
+
+\begin{figure}
+\label{fig:block_model_2}
+\includegraphics[scale=.25]{block_model_4}
+\caption{$n_D$ against $ p_{in} $ for systems of $N=200 $, $p_{out} = 0.1 $, and for different values of $N_{clusters} $. The curves are averaged over 100 realizations.}
+\end{figure}
+
+Until now, we have considered all batteries to be equal, i.e they have the same capacity and charge/discharge rate.
+Although simple, this assumption might not be verified in practice where different types of batteries exist.
+Making a cost-oriented study where the location os various types of batteries is studied such that the tradeoff between costs and reliability is out of the scope of this paper.
+Nevertheless, we studied how the number of drivers behaves when the batteries have different capcities.
+More precisely, we draw the capacities of the batteries from a normal distribution $ \mathcal{N}( \mu_{\lambda}, \sigma_{\lambda} ) $ and keep a simple erdos-renyi topology for the power grid.
+Note that each node is assigned a capacity regardless of its caracteristics (degree, betweeness, and so on...).
+In figure 6, we plot $n_D$ as a function of the relative standard deviation of the battery capacity distribution $ \sigma_{\lambda} / \mu_{\lambda} $.
+We see that $n_D$ increases with the variance of the capacity distribution for all three gramian based metrics considered.
+Moreover, this rise is abrupt since $n_D$ goes from $ 60 \%$ for $\frac{\sigma_{\lambda}}{\mu_{\lambda}} \sim 0.15 $ to $100 \% $ for $\frac{\sigma_{\lambda}}{\mu_{\lambda}} \sim 0.27 $.
+
+\begin{figure}
+\label{fig:batteries_variance}
+\includegraphics[scale=.37]{batteries_variance}
+\caption{$n_D$ against the relative standard deviation of the capacity distribution of the batteries $ \frac{ \sigma_{ \lambda } }{ \mu_{ \lambda} } $.
+The dots are averages over 100 realizations and the curves are obtained using a Savitzky-Golay filter, each color shows the results for a given metric.
+The graphs are erdos-renyi with $N=100$ and $p=0.3$, and $\mu_{\lambda} = 100 $ units.
+}
+\end{figure}
+
+Real power grids topologies are known to be far from random such that using the erdos-renyi model probably do not give a realistic insight into the control of these systems.
+Depending on the voltage level
+
+In this section, we provide some results in order to illustrate the work described above.
+We use a small 10 prosumers example, connected with a random topology.
+The line maximum capacities are selected  according to a normal distribution $\mathcal{N}(\mu_l,\sigma_l)$.
+The power distribution is chosen as a zero mean normal distribution $ \mathcal{N}(0,\sigma_p) $, and the samples should satisfy the constraints \ref{synchro_constraint} and \ref{zero_sum_constraint}.
+Obviously, $\mu_l$, $\sigma_l$, and $\sigma_p$ should be chosen such that, without any perturbation, the lines are not already overloaded.
 

### 2016-02-15 Article.tex
intro modifs

@@
 Modernizing the power grids and increasing the share of renewables in the production are two substantial objectives of the energetic transition.
-Because of parallel progresses in communication, data management, and storage, the upcoming emergence of a power grid "2.0", often called smart grid, appears as a cross-disciplinary challenge of the 21st century \cite{Ramchurn}.
+Because of progresses in communication, data management, and storage, the upcoming emergence of a power grid "2.0", often called smart grid, appears as a cross-disciplinary challenge of the 21st century \cite{Ramchurn}.
 
@@
 In this paper, we explore the use of storage in a network of prosumers whose power outputs are susceptible to change due to various external causes.
-More precisely, we are interested in the "best" locations for the storage equipments such that we are able to maintain stability when  perturbations occur.
+
+%More precisely, we are interested in the "best" locations for the storage equipments such that we are able to maintain stability when  perturbations occur.
 Since all nodes can switch from generator to load and vice versa, the optimal location is susceptible to change with the prosumers power output.
-Our goal is then to find the one that, on average, provide the best results.
+Our goal is then to find the one that, on average, provides the best results.
 
-Since charging and discharging a battery is costly, we are particularly interested in the energy necessary to control the system when perturbations occur.
+%Since charging and discharging a battery is costly, we are particularly interested in the energy necessary to control the system when perturbations occur.
 The notion of performance of a given set of locations for using storage would then be inversely proportional to the energy used.
@@
 
- In this paper, we model the prosumer network as a coupled oscillator network.
-Inspired by \cite{Filatrella2008}, we use a second order Kuramoto model that can be seen as a simplified version of the underlying power grid dynamics (see Section \ref{sec:The_Model}).
-We then use control theory \cite{Liu2015} in order to model the impact of the storage equipments on the network (see Section \ref{sec:control_and_submodularity}).
-Using the relationships between submodular set functions and the Gramian matrix \cite{Summers2014}, we use a modified version of a greedy algorithm \cite{Minoux} in order to find the smallest locations set that minimizes the average control energy while satisfying all the constraints (see section \ref{sec:Control_prosumer}).
-Finally, we show some results in Section \ref{sec:Results}.
+We model this network as a coupled oscillator network, where each node is represented as an oscillator with a phase angle and a natural frequency that are impacted by the ones of their neighbors.
+Since the work of Kuramoto [], networks of coupled oscillators have been used to model various phenomenons such as fireflies, ex2, or power grids.
+Indeed, it is well-known that synchronization of all rotating elements to a common frequency is necessary within a power grid for stable operation.
+Inspired by \cite{Filatrella2008}, we use a second order Kuramoto model in order to obtain a simplified dynamics for the power grid (see Section \ref{sec:The_Model}).
+
+We will see that under some conditions, the system synchronizes to a common frequency.
+However, this synchronized state is suceptible to break if perturbations occur.
+If, for example, the production of some generator decreases while the load remains the same, the imbalance between production and consumption might drive the system to unstable conditions.
+These kinds of perturbations are very likely in a scenario with high penetration of DER.
+Multiple control techniques, with different time scales, exits in order to cope with such events.
+Among others, it has been argued that storage and vehicle to grid (V2G) technology could be used for frequency regulation.
+If a sudden drop or raise of the frequency due to an imbalance is detected, some devices are charged or discharged as to restore synchrony.
+This approach poses a few questions such as the location of the storage devices, or the quantities of power that should be injected or absorbed.
+Since charging and discharging a battery is costly, we are particularly interested in the energy necessary to restaure synchrony.
+Moreover, in a prosumer scenario where generators and loads tend not to be fixed, we would like to find a subset of locations for installing storage such that the energy is low on average.
+
+We will see that these storage actions can be interpreted as an additional term in the oscillators dynamics for a subset of nodes.
+We propose here to use complex network control theory \cite{Liu2015} in order to model the impact of the storage equipments on the network dynamics (see Section \ref{sec:control_and_submodularity}).
+Placing storage at some node can thus be seen as selecting this node in the driver set.
+Selecting the drivers that, on average, will control the network with low energy is a hard problem.
+Nevertheless, recent studies showed the existence of relationships between controllability and submodular set functions \cite{Summers2014}.
+These functions, that we introduce in more depth in section X, have the interesting property of having diminishing returns as we increase the size of the set considered.
+Although finding the optimal set is NP-hard, \cite{Minoux} proposed a simple greedy algorithm with a worst case guaranty.
+
+Since the system under study is a power grid which lines and batteries have finite capacities, we have further constraints on the control signals.
+Obviously, no line can be overloaded while trying to restaure synchrony.
+Likewise, no battery can inject power when empty, or charge/discharge quicker than some maximum rate.
+We propose here to incorporate these constraints in the search of the driver set (see section \ref{sec:Control_prosumer}).
+
+This paper is organized as follows, section \ref{sec:The_Model} presents the oscillator model used for simplifying the grid's dynamics.
+Section 3 introduces submodular functions and their maximization.
+Section 4 gives basic notions of control theory and links controllability to submodular set functions.
+In section \ref{sec:Control_prosumer} we develop the constraints relative to the system and the batteries.
+Finally, in section \ref{sec:Results}, we show some results.
 
@@
 
-For a direct illustrative application, we chose a small system of 15 nodes connected with an erdos-renyi topology.
+For a direct illustrative application, we chose a small system of 10 nodes connected with an erdos-renyi topology.
 The power distribution and the capacities of the lines are selected randomly such that the system is at equilibrium.
@@
 In other words, if we select the drivers randomly, we expect to need, on average, more energy to control the system.
-In figure 2, we draw 200 example systems of 15 nodes with random topologies, power distributions, and line capacities.
-These systems are initially at equilibrium and are perturbed as explained previously.
-For each system we find the driver set S using algorithm (X) and we draw a random driver set $S'$ such that $S$ and $S'$ have the same size, and such that $S'$ also provide full control of the system.
+In figure 2, we draw $10^4$ example systems of 50 nodes with random scale-free topologies (Barabasi-Albert model is used), and random power distributions and line capacities.
+For each system we select a random number of drivers $N_D \sim \mathcal{U}(1,N_{nodes}) $ and we find two driver sets of size $N_D $.
+One is chosen randomly and the other is found by greedy optimization of a gramian based metric.
+For both sets, we draw a random initial state $Y_i$ and a random final state $Y_f$, and we compute the control energy required for driving the system from $Y_i$ to $Y_f$.
+On figure \ref{fig:densities} we plot each system as a dot with coordinates $(n_D = N_D / N_{nodes}, log(\mathcal{E}))$.
+The color indicates what method has been used, and the lines show the average accross $n_D$.
+As visible, both curve decreases as the number of drivers increases, meaning that, as the number of drivers grows, the average control energy tends to decrease.
+Furthermore, we tend to use, as exected, less energy when the drivers are carefully selected.
+Note that this difference tends to zero when $n_D$ tend to one, because almost all nodes are then selected, yielding little flexibility for optimization.
+
+%For each system we find the driver set S using algorithm (X) and we draw a random driver set $S'$ such that $S$ and $S'$ have the same size, and such that $S'$ also provide full control of the system.
 We then compute the control signals and the corresponding control energies $\mathcal{E}_S$ and $\mathcal{E}_{S'}$ for both $S$ and $S'$.
@@
 \label{fig:densities}
-\includegraphics[scale=.3]{rr}
-\caption{test}
+\includegraphics[scale=.35]{energy}
+\caption{Control energy $log(\mathcal{E})$ against size of control set $n_D$ for random and optimized driver sets.
+Topologies are drawn from a Barabasi-Albert model with $ N_{nodes} = 50 $, power and line capacity distributions are random.
+For both case, $10^4$ instances are considered and the results are displayed as dots.
+The lines shows the average for both case.
+}
 \end{figure}
@@
 \begin{itemize}
-\item Level 1 : full control
-\item Level 2 : full control + line capacities constraints + battery constraints (see algorithm X)
+\item Level 1 : full control : the system can be moved from any point to any other point of the state space. We do not consider constraints on the energy which can tend to infinity. 
+\item Level 2 : full control + line capacities constraints + battery constraints (see algorithm X).
+This level is much more demanding since we impose the ability to move the system from any point to any other point, without overloading any line or breaking any battery limits in the process.
 \end{itemize}
@@
 
-Figure 4 shows how the number of drivers evolves with the number of clusters for different values of $ p_{in} $ when $ p_{out} $ is fixed ($p_{out} =0.1$).
+Figure 4 displays how the number of drivers evolves with the number of clusters for different values of $ p_{in} $ when $ p_{out} $ is fixed ($p_{out} =0.1$).
 For small values of $ p_{in} \sim p_{out} $ clusters are poorly marked, and the connectivity is low.
@@
 Although simple, this assumption might not be verified in practice where different types of batteries exist.
-Making a cost-oriented study where the location os various types of batteries is studied such that the tradeoff between costs and reliability is out of the scope of this paper.
-Nevertheless, we studied how the number of drivers behaves when the batteries have different capcities.
+Optimizing the locations of different types of storage as to optimize the cost / reliability tradeoff is out of the scope of this paper.
+Nevertheless, we studied how the number of drivers behaves when the batteries have different capacities.
 More precisely, we draw the capacities of the batteries from a normal distribution $ \mathcal{N}( \mu_{\lambda}, \sigma_{\lambda} ) $ and keep a simple erdos-renyi topology for the power grid.

### 2016-02-17 Article.tex
update

@@
 
-Modernizing the power grids and increasing the share of renewables in the production are two substantial objectives of the energetic transition.
+Modernizing the power grid and increasing the share of renewables in the production are substantial objectives of the energetic transition.
 Because of progresses in communication, data management, and storage, the upcoming emergence of a power grid "2.0", often called smart grid, appears as a cross-disciplinary challenge of the 21st century \cite{Ramchurn}.
 
-It is believed that these new ways of producing and managing energy, will also impact how end users consume.
-More precisely, the consumer   of the future is expected to be more involved in the system operation.
-Using dynamic signals (real time electricity prices for instance), load curves could be shaped to some extend as to agree with the production conditions of the main grid.
-Furthermore, the emergence of renewable distributed energy resources (DER) in the distribution networks, possibly hold by individuals,
-will completely modify the today end user agent model.
+The today centralized top-down architecture is set to evolve to more distributed and bi-directional systems.
+Furthermore, the emergence of renewable and stochastic distributed energy resources (DER) in the distribution networks will require more flexibility.
+All these evolutions might change deeply the ways end users perceive and consume electricity.
+The consumer of the future, while seeking maximum utility/profit tradeoffs, is expected to be more involved in the system operation.
+Using dynamic signals such as real time electricity prices for instance, load curves could be shaped to some extend as to agree with the production conditions of the grid.
+Within these complex systems with numerous agents, it is assumed that multiple aggregation levels will be required in order to organize and optimize communication.
+Agents responsible for generation and load portfolio managment, i.e prosumers \cite{Rathnayaka2012}, will provide services (generation, load shedding, frequency regulation) against remuneration.
+Insure grid stability within this uncertain context appears as a complex task.
 
-These new agents that can both consume and produce are often called prosumers \cite{Rathnayaka2012}.
-This decentralized production is believed to lead to more resilient systems than today centralized power grid.
-Nonetheless, the difficulty of dealing with decentralized production is coupled with the high uncertainty of renewables.
-Ensure system stability in these conditions appears thus as a complex task.
-
-Relying on the ability of the smart grid to obtain, communicate and analyse data, several solutions could  improve the stability of such complex systems.
-Among these solutions, demand side management and energy storage are two of the most popular ones.
+Several studies highlighted the need for increasing the storage capacity of the system.
+Some of them even consider electric vehicles as moving capacities that could be used for frequency regulation.
+In the case of fixed storage devices, the locations where they should be installed is an important question since it impacts the performances of the system.
+In a very static and top-down architecture, where loads and generators are fixed, this question is far from trivial.
+But in a smart grid scenario with bi-directional flows and agents that produce and consume depending on weather conditions, this becomes even more challenging.
 In this paper, we explore the use of storage in a network of prosumers whose power outputs are susceptible to change due to various external causes.
@@
 
-We model this network as a coupled oscillator network, where each node is represented as an oscillator with a phase angle and a natural frequency that are impacted by the ones of their neighbors.
+More precisely, we model the power grid as a coupled oscillator network, where each node is represented as an oscillator with a phase angle and a natural frequency that are impacted by the ones of their neighbors.
 Since the work of Kuramoto [], networks of coupled oscillators have been used to model various phenomenons such as fireflies, ex2, or power grids.
@@
 
-We will see that under some conditions, the system synchronizes to a common frequency.
+We will see that under some conditions, the system synchronizes to a common frequency $\Omega$.
 However, this synchronized state is suceptible to break if perturbations occur.
-If, for example, the production of some generator decreases while the load remains the same, the imbalance between production and consumption might drive the system to unstable conditions.
+If, for example, the production of some generator decreases while the load remains the same, the imbalance between production and consumption might steer the frequency away from $\Omega$ and drive the system to unstable conditions.
 These kinds of perturbations are very likely in a scenario with high penetration of DER.
@@
 We propose here to use complex network control theory \cite{Liu2015} in order to model the impact of the storage equipments on the network dynamics (see Section \ref{sec:control_and_submodularity}).
-Placing storage at some node can thus be seen as selecting this node in the driver set.
-Selecting the drivers that, on average, will control the network with low energy is a hard problem.
-Nevertheless, recent studies showed the existence of relationships between controllability and submodular set functions \cite{Summers2014}.
+Placing storage at some node can thus be seen as selecting this node in the driver set, such that finding the locations for the storage devices could be related to finding the driver set of the network.
+Nevertheless, selecting the drivers that, on average, will control the network with low energy is a hard problem.
+Recent studies showed though the existence of relationships between controllability and submodular set functions \cite{Summers2014}.
 These functions, that we introduce in more depth in section X, have the interesting property of having diminishing returns as we increase the size of the set considered.
@@
 The condition for the power transmission between i and j is that the two devices do not operate in phase : The phase difference between i and j is : $ \delta_j(t) - \delta_i(t) = \Omega t + \theta_j(t) - \Omega t - \theta_i(t) = \theta_j(t) - \theta_i(t) $.
-
-The transmitted power along the line can thus be written as :
+The transmitted power along the line can be written as :
 
@@
 \begin{equation}
+\label{eq:without_simplification}
 \begin{array}{lll}
@@
 
-These non linear equations are obviously not very practical.
-We now use some approximations based on the fact that we consider only small deviations from the main frequency : $ \dot{ \delta}_i \sim \Omega $ which means that $ \omega_i = \dot{\theta_i} << \Omega $, such that the squared term $ K_D \dot{\theta}_i^2 $ can be neglected.
-Moreover, we assume that the rate at which energy is stored in the kinetic term is much less of the rate at which energy is dissipated by friction, a common condition for mechanical systems : $ \ddot{ \theta }_i  << \frac{2 K_D}{I} $
-
-The equation can thus be simplified to :
+We now use simplifications based on the fact that we consider small deviations from the main frequency : $ \dot{ \delta}_i \sim \Omega $ which means that $ \omega_i = \dot{\theta_i} << \Omega $, such that the squared term $ K_D \dot{\theta}_i^2 $ can be neglected.
+Moreover, we assume that the rate at which energy is stored in the kinetic term is much less of the rate at which energy is dissipated by friction : $ \ddot{ \theta }_i  << \frac{2 K_D}{I} $ (see \cite{Filatrella2008} for more details).
+Equation \ref{eq:without_simplification} becomes :
 
@@
 \begin{equation}
+\label{eq:with_simplification}
 \ddot{\theta} = \Psi - \alpha \dot{\theta} - KL\theta
@@
 
-This is a continuous time second order linear system of $ N $ equations.
+Equation \ref{eq:with_simplification} is a continuous time second order linear system of $ N $ equations.
 We first transform this into a continuous time first order linear system of $ 2 N $ equations by introducing the following vector : $ X = \left( \begin{array}{c} \theta \\ \dot{\theta} \end{array} \right)$ :
@@
 \begin{equation}
+\label{eq:final_equation}
  Y(t+\Delta t) = A Y(t)
@@
 
-This is a linear discrete time system of $ 2 N + 1 $ equations. Note that the transition matrix A encodes all the system parameters, topology, and power distribution.
+Equation \ref{eq:final_equation} is a linear discrete time system of $ 2 N + 1 $ equations. Note that the transition matrix A encodes all the system parameters, topology, and power distribution.
 
@@
 
+\begin{figure}
+	\label{fig:submodular}
+	\includegraphics[scale=.5]{submodular}
+	\caption{Submodularity example with sensor placements.}
+\end{figure}
+
 This basically means that submodular functions exhibit a diminishing return property which makes them particularly interesting for optimization.
+A very intuitive example is the optimum placement of sensors in an area (see figure \ref{fig:submodular}).
+Sensors can be placed on a grid of locations and function F computes the surface of the space that is being sensed.
+Note first that adding a new sensor i to the current set S cannot decrease the value of F : $F(S+\{i\})>=F(S)$.
+Furthermore, if we add sensor i to a small set $S_1$ we tend to get larger improvements than if we add i to a larger set $S_2 \supset S_1$.
 
@@
 Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k) \cite{Krause2014}.
+This heuristic starts with a set S (possibly empty) and add the element i that exhibits the highest marginal gain : $ F(S \cup \{i\}) \geq F(S \cup \{j\})\ \forall j $.
 
+For a ground set of N elements, this heuristic computes F $ \frac{k(2 N - k + 1)}{2} $ times.
 Since the evaluation of F can be costly, a well-known lazy-greedy variation has been proposed by \cite{Minoux}.
 This smart implementation uses the submodular structure of the marginal gains in order to reduce the number of calls to F.
-This can result in speedups of several order of magnitudes.
+This requires to maintain a sorted table of marginal gains for all elements.
+When looking for new element i to add to set S, the top one is selected and the new marginal gain $F(S\cup \{§i\}) - F(S)$ is computed.
+If this gain is larger than the  gain of the second element in the table, then i is added to S.
+Otherwise i is inserted back in the table with its updated gain and the same treatment is applied to the element that is now on the top of the table.
+Because of the submodularity of F, this method performs as well as the original one, but can result in speedups of several order of magnitudes.
 
@@
 We assume the control of a subset of the nodes that we call driver nodes.
-In our particular power grid situation, a driver node can be seen as a prosumer where some additional equipment (battery, backup generator,...) will be used as to inject or absorb energy from the system.
-This control assumption modifies the system dynamic :
+In our particular power grid situation, a driver node can be seen as a prosumer where some additional storage equipment is used to inject or absorb energy from the system.
+The dynamics of equation \ref{eq:final_equation} becomes :
 
@@
 Let $ \mathcal{E} = \int_{t_0}^{t_0+T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system.
-
 It can be shown \cite{Liu2015} that the control input sequence that minimizes the control energy can be written as :
@@
 Therefore, we concentrate on  the Gramian as a way to quantify the average controllability.
-It can indeed be verified that the Gramian does not depend on $\Psi$.
 
-A key point to design a low consumption control is thus to understand the relationships between the control node set and the Gramian matrix.
+Since the control energy is related to $ W^{-1} $, small energies are obtained from "large" W.
 In [], the authors introduce several Gramian-based metrics that have concrete meanings in terms of the system controllability.

### 2016-02-22 figure_3/plot_figure_3.py
big figures update

@@ -1,0 +1,47 @@
+import json
+import numpy as np
+import matplotlib.pyplot as plt
+import random
+import math
+import json
+import numpy as np
+import matplotlib.pyplot as plt
+import random
+import math
+from scipy.signal import savgol_filter
+
+
+with open('./test_OPT.json', 'r') as f:
+    OPT = json.load(f)
+with open('./test_RAND.json', 'r') as f:
+    RAND = json.load(f)
+
+OPT = np.nan_to_num(np.array(OPT) )
+RAND = np.nan_to_num( np.array(RAND) )
+
+with open('./test_OPT.json', 'r') as f:
+    OPT = json.load(f)
+OPT = np.nan_to_num(np.array(OPT) )
+
+N_nodes = 50
+
+MEANS = [ np.mean( RAND[RAND[:,0]==i][:,1] ) for i in range(1,N_nodes) ]
+
+MEANS2 = [ np.mean( OPT[OPT[:,0]==i][:,1] ) for i in range(1,N_nodes) ]
+
+
+MEANS = savgol_filter(MEANS, 5, 3)
+MEANS2 = savgol_filter(MEANS2, 5, 3)
+
+fig = plt.figure(figsize=(10,10))
+plt.scatter( 1./N_nodes * RAND[:,0], RAND[:,1], s= 60, color = '#FD6AF2', alpha = .3 )
+plt.scatter( 1./N_nodes * OPT[:,0], OPT[:,1], s= 60, color = '#FDAC86', alpha = .2 )
+plt.plot( 1./N_nodes * np.arange(1,N_nodes), MEANS, linewidth = 4, color = '#C703B8', label = 'random' )
+plt.plot( 1./N_nodes * np.arange(1,N_nodes), MEANS2, linewidth = 4, color = 'r', label = 'greedy optimization' )
+plt.xlim((0.4,1))
+plt.ylim((5,22))
+plt.xlabel(r'$n_D$', fontsize = 20)
+plt.ylabel(r'$log(\mathcal{E})$', fontsize = 20)
+plt.legend(loc=3, fontsize = 15)
+plt.grid()
+plt.savefig('./figure_3.pdf', dpi = 300)

### 2016-03-01 Article.tex
big figures update
update

@@
 	\label{fig:submodular}
-	\includegraphics[scale=.5]{submodular}
+	\begin{center}
+	\includegraphics[scale=.3]{./figure_1/figure_1}
 	\caption{Submodularity example with sensor placements.}
+	\end{center}
 \end{figure}
@@
 \begin{equation}
+\label{eq:final_with_control}
  Y(t+\Delta t) = A Y(t) + B u(t) 
@@
 The system is said to be controllable in T steps if it can be steered from any initial state $ Y_0 $ to any final state $ Y_f $ through a sequence of control inputs $ u(t) $.
-If the system is controllable, there might be more than one sequence of control inputs that could achieve such a thing.
-Among all these possibilities, we are interested in the one that requires the minimum amount of energy \cite{Yan2012}.
+
+Controllability in complex network became popular with the well-known structural controllability framework introduced by Lin \cite{1100557}.
+This approach relates the search of the minimum driver set to a maximum matching in a bipartite graph, which can be done with the Hopcroft-Karp algorithm in $ O(|E|\sqrt{|V|}) $ in the worst case.
+This requires that the system $(A,B)$ is a structured system, meaning that all elements in A and B are either fixed zeros or independant free parameters.
+In such a situation, structural control is extremely powerfull since it enables us to discuss a network's controllability relatively quickly even if we do not know the exact weights of the edges.
+Note however that the independance of the parameters is a key assumption.
+If A is the adjacency matrix of an undirected network for instance, independance is not true since A is symetric, and one must use different tools to study controllability \cite{Yuan2014}.
+
+On the other hand, if the system is controllable, there might be more than one sequence of control inputs $u(t)$ that could drive it from $Y_0$ to $Y_f$.
+Among all these possibilities, optimal control is devoted to find the sequence that minimize some cost function J that can depend on the sates of the system $Y(t)$, the control inputs $u(t)$, and the final state $Y(T)$.
+As explained above, we are interested in the sequence that requires the minimum amount of control energy \cite{Yan2012}.
 Let $ \mathcal{E} = \int_{t_0}^{t_0+T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system.
@@
 \begin{equation}
+\label{eq:ustar}
  u^{\star}(t) = B^T(A^T)^{T-t-1}W^{-1}(T)\nu(T)
@@
 Since the control energy is related to $ W^{-1} $, small energies are obtained from "large" W.
-In [], the authors introduce several Gramian-based metrics that have concrete meanings in terms of the system controllability.
+In \cite{Summers2014}, the authors introduce several Gramian-based metrics that have concrete meanings in terms of the system controllability.
 For example, the trace of the inverse Gramian ($ Tr[ W^{-1}(T)] $) quantifies the average control energy necessary to move the system around the state space.
@@
 Using the Gramian-based metric $ \mathcal{M} $, let $ F(S) =\mathcal{M}(W_S(T))$, where $ W_S(T) $  is the Gramian matrix based on the driver node set S.
-[] showed that for some metrics, F is a submodular set function.
+\cite{Summers2014} showed that for some metrics, F is a submodular set function.
 This means that, for these metrics, we can use the greedy heuristic discussed above in order to maximize F.
@@
 
- Considering the dynamic of equation (), we notice that the Gramian based on the transition matrix A is never invertible, meaning that the system should not be controllable.
+ Considering the dynamic of equation \ref{eq:final_with_control}, we notice that the Gramian based on the transition matrix A is never invertible, meaning that the system should not be controllable.
 It is easy to see that, indeed, we do not have full control because of the last row of the matrix.
@@
 \begin{algorithm}
-	\label{algo_1}
-
         \begin{algorithmic}
@@
         \end{algorithmic}
+        \caption{Greedy submodular optimization with constraints}
+        \label{algo_1}
 \end{algorithm}
@@
 
-For a direct illustrative application, we chose a small system of 10 nodes connected with an erdos-renyi topology.
+We first consider a small example of 10 nodes connected with an erdos-renyi topology.
 The power distribution and the capacities of the lines are selected randomly such that the system is at equilibrium.
 Therefore, the production matches the consumption, no line is overloaded and all nodes are synchronized to the main frequency.
-We use the greedy algorithm (X) based on the trace of the inverse Gramian as to select the driver set $S$.
+We use the greedy algorithm \ref{algo_1} based on the trace of the inverse Gramian as to select the driver set $S$.
 At a given time $ t_k $, we introduce a small perturbation in the form of a mismatch between the production and consumption : $P_i(t_k) = P_i(t_{k-1}) + \delta P$ for some random node i.
 Consequently, the frequencies $ \dot{ \theta }_i \forall i $ of the oscillators start to deviate from the synchronized state.
-At time $ t_k + d $, we apply to the drivers the optimal control of equation (X) such that the system is brought to the synchronized state at time $ t_k + d + T $, where T is the control time.
+At time $ t_k + d $, we apply to the drivers the optimal control of equation \ref{eq:ustar} such that the system is brought to the synchronized state at time $ t_k + d + T $, where T is the control time.
 At this time, if the mismatch between production and consumption is still active, we maintain synchrony by using optimal control with $T=1$.
-Figure 1 shows the evolution of a subset (for clarity) of the nodes' frequencies during the control phase.
+Figure \ref{fig:frequecies} shows the evolution of the nodes' frequencies during the control phase.
 
 \begin{figure}
-\label{fig:frequecies}
-\includegraphics[scale=.38]{frequencies}
+\includegraphics[scale=.38]{./figure_2/figure_2.pdf}
 \caption{Time series of the N frequencies $\omega_i$ (phase angles are not shown for clarity).
@@
 Note that, if the perturbation is still active at this point, the system still needs control in order to stay synchronized (batteries should still compensate the power imbalance)}
+\label{fig:frequecies}
 \end{figure}
@@
 In other words, if we select the drivers randomly, we expect to need, on average, more energy to control the system.
-In figure 2, we draw $10^4$ example systems of 50 nodes with random scale-free topologies (Barabasi-Albert model is used), and random power distributions and line capacities.
+In figure \ref{fig:densities}, we draw $10^4$ example systems of 50 nodes with random scale-free topologies (Barabasi-Albert model is used), and random power distributions and line capacities.
 For each system we select a random number of drivers $N_D \sim \mathcal{U}(1,N_{nodes}) $ and we find two driver sets of size $N_D $.
-One is chosen randomly and the other is found by greedy optimization of a gramian based metric.
+One is chosen randomly and the other is found with algorithm \ref{algo_1}.
 For both sets, we draw a random initial state $Y_i$ and a random final state $Y_f$, and we compute the control energy required for driving the system from $Y_i$ to $Y_f$.
@@
 The color indicates what method has been used, and the lines show the average accross $n_D$.
-As visible, both curve decreases as the number of drivers increases, meaning that, as the number of drivers grows, the average control energy tends to decrease.
+As visible, both curves decrease as the number of drivers increases, meaning that, as the number of drivers grows, the average control energy tends to decrease.
 Furthermore, we tend to use, as exected, less energy when the drivers are carefully selected.
-Note that this difference tends to zero when $n_D$ tend to one, because almost all nodes are then selected, yielding little flexibility for optimization.
+Note that this difference tends to zero when $n_D$ tends to one, because almost all nodes are then selected, yielding little flexibility for optimization.
 
@@
 \begin{figure}
-\label{fig:densities}
-\includegraphics[scale=.35]{energy}
+\includegraphics[scale=.35]{./figure_3/figure_3}
 \caption{Control energy $log(\mathcal{E})$ against size of control set $n_D$ for random and optimized driver sets.
@@
 }
+\label{fig:densities}
 \end{figure}
@@
 We consider first the case of an erdos renyi topology with probability of connection p.
-We show on Figure 3 how the size of the driver set evolves with p for different Gramian based metrics and for two levels of constraints :
+We show on Figure \ref{fig:erdos_renyi} how the size of the driver set evolves with p for different Gramian based metrics and for two levels of constraints :
 
@@
 \item Level 1 : full control : the system can be moved from any point to any other point of the state space. We do not consider constraints on the energy which can tend to infinity. 
-\item Level 2 : full control + line capacities constraints + battery constraints (see algorithm X).
+\item Level 2 : full control + line capacities constraints + battery constraints (see algorithm \ref{algo_1}).
 This level is much more demanding since we impose the ability to move the system from any point to any other point, without overloading any line or breaking any battery limits in the process.
@@
 \begin{figure}
-\label{fig:erdos_renyi}
-\includegraphics[scale=.3]{graphique_1}
+\includegraphics[scale=.25]{./figure_4/figure_4}
 \caption{ $ n_D $ against link probability p for erdos-renyi topologies ($ N=100 $).
@@
 The bottom three curves exhibit the results for the same metrics when only the full controllability constraint is considered (see Level 1 in the main text).}
+\label{fig:erdos_renyi}
 \end{figure}
@@
 \begin{figure}
+\includegraphics[scale=.25]{./figure_5/figure_5}
+\caption{$n_D$ against $N_{clusters} $ for different values of $ p_{in} $ and $ p_{out} = 0.1 $. $ N=200 $ and curves are averaged over 100 realizations.}
 \label{fig:block_model_1}
-\includegraphics[scale=.25]{block_model_3}
-\caption{$n_D$ against $N_{clusters} $ for different values of $ p_{in} $ and $ p_{out} = 0.1 $. $ N=200 $ and curves are averaged over 100 realizations.}
 \end{figure}
 
-Figure 4 displays how the number of drivers evolves with the number of clusters for different values of $ p_{in} $ when $ p_{out} $ is fixed ($p_{out} =0.1$).
+Figure \ref{fig:block_model_1} displays how the number of drivers evolves with the number of clusters for different values of $ p_{in} $ when $ p_{out} $ is fixed ($p_{out} =0.1$).
 For small values of $ p_{in} \sim p_{out} $ clusters are poorly marked, and the connectivity is low.
@@
 As the number of clusters grows, the global connectivity increases rapidely, such that the number of drivers also rises.
-We show this behavior in more details in figure 5, where the number of drivers is plotted against $ p_{in} $ for $N=200$ and $N_{clusters} \in \{2,4,5\} $.
+We show this behavior in more details in figure \ref{fig:block_model_2}, where the number of drivers is plotted against $ p_{in} $ for $N=200$ and $N_{clusters} \in \{2,4,5,10\} $.
 For $N_{clusters} = 2 $ and $ p_{in} = 0.1 $ the number of drivers is large since the graph is poorly connected.
-When $p_{in} $ increases, $n_D$ decreases until a minimum value $n_D \sim 0.64 $ at $ p_{in} \sim 0.6 $.
+When $p_{in} $ increases, $n_D$ decreases until a minimum value $n_D \sim 0.6 $ at $ p_{in} \sim 0.25 $.
 After this point, $n_D$ augments with $p_{in}$.
-For $N_{clusters} = 5 $, we do not see this behavior : $ n_D $ decreases continuously as $p_{in} $ increases as expected from the curves of figure 4.
+For $N_{clusters} = 10 $, we do not see this behavior : $ n_D $ decreases continuously as $p_{in} $ increases as expected from the curves of figure \ref{fig:block_model_1}.
 
 \begin{figure}
+\includegraphics[scale=.25]{./figure_6/figure_6}
+\caption{$n_D$ against $ p_{in} $ for systems of $N=200 $, $p_{out} = 0.1 $, and for different values of $N_{clusters} $. The curves are averaged over 100 realizations.}
 \label{fig:block_model_2}
-\includegraphics[scale=.25]{block_model_4}
-\caption{$n_D$ against $ p_{in} $ for systems of $N=200 $, $p_{out} = 0.1 $, and for different values of $N_{clusters} $. The curves are averaged over 100 realizations.}
 \end{figure}
@@
 \begin{figure}
-\label{fig:batteries_variance}
-\includegraphics[scale=.37]{batteries_variance}
+\includegraphics[scale=.37]{./figure_7/figure_7}
 \caption{$n_D$ against the relative standard deviation of the capacity distribution of the batteries $ \frac{ \sigma_{ \lambda } }{ \mu_{ \lambda} } $.
@@
 }
+\label{fig:batteries_variance}
 \end{figure}
 
-Real power grids topologies are known to be far from random such that using the erdos-renyi model probably do not give a realistic insight into the control of these systems.
-Depending on the voltage level
+\begin{figure}
+\includegraphics[scale=.25]{power_grid}
+\caption{European transmission power grid \cite{jensen_2015_35177}. Nodes are colored according to their county.}
+\label{fig:network}
+\end{figure}
 
-In this section, we provide some results in order to illustrate the work described above.
+Since real power grids topologies are known to be far from random, we consider here, as a real case, the european transmission power grid which topology was obtained from \cite{jensen_2015_35177}.
+The network contains 1494 nodes and 2156 edges and spans 25 european countries.
+A representation of this graph can be found in figure \ref{fig:network} where the nodes are colored according to their country.
+We select the power distribution and the line capacities randomly, and use the method described above.
+
+We use the trace of the Gramian matrix for the optimization.
+The top panel of figure \ref{fig:real_case} shows that the rank of the Gramian increases with the size of the controllers until reaching full rank ($rank[W_S] = 2 N_{nodes}$) for $N_D=1244$.
+Above this point, we compute the average control energy needed to drive the system from an initial random state to a random final state.
+The curves on the bottom panel of figure \ref{fig:real_case} show the results for the optimized driver sets and for random driver sets.
+As can be seen on figure \ref{fig:real_case}, as the size of the driver set increases, the control energy tend to decrease.
+But, we need less energy with the optimized driver set than with randomly sampled sets.
+
+\begin{figure}
+\includegraphics[scale=.4]{figs2}
+\caption{Gramian rank and control energy evolution for european transmission power grid (see figure \ref{fig:network}).}
+\label{fig:real_case}
+\end{figure}
+%In this section, we provide some results in order to illustrate the work described above.
 We use a small 10 prosumers example, connected with a random topology.
@@
 
-Once the synchronized state is reached, we apply a perturbation to the power distribution such that there is a mismatch between production and demand (constraint \ref{zero_sum_constraint} is not satisfied).
+%Once the synchronized state is reached, we apply a perturbation to the power distribution such that there is a mismatch between production and demand (constraint \ref{zero_sum_constraint} is not satisfied).
 We denote by $d$ the delay between the perturbation and the start of the control phase (see figure \ref{fig:frequecies}).
@@
 
- If the driver set is fixed, we expect that the higher $ d$, the higher the control energy required to drive the system back to the synchronized state.
+%If the driver set is fixed, we expect that the higher $ d$, the higher the control energy required to drive the system back to the synchronized state.
 Eventually, if $d$ is too large, the minimum energy control inputs will result in some constraints being broken (battery levels, line flows, and so on...).
 
-\begin{figure}
-\label{fig:energy_vs_delay}
-\includegraphics[scale=.38]{energy_vs_delay2}
-\caption{ Control energy used in order to bring the system back to the synchronized state when the delay between perturbation and control increases, while the driver node set is fixed.
-The blue dotted curve shows the results for $ S_k^{\star} $ our algorithm \ref{algo_1}.
-The red line shows (average and standard deviation) the results when the driver set is sampled uniformly from all possible sets of the size $ k $.}
-\end{figure}
-
-As visible on figure \ref{fig:energy_vs_delay}, when d increases, the system deviates more from the equilibrium and this results in more energy being used to steer it back to synchrony.
+%As visible on figure \ref{fig:energy_vs_delay}, when d increases, the system deviates more from the equilibrium and this results in more energy being used to steer it back to synchrony.
 We also investigate here how the driver node set selected with algorithm \ref{algo_1} performs against other driver sets of the same size.
@@
 \label{sec:Conclusion}
-C'est fini !
+In this paper, we considered the case of networks of prosumers which can behave as generators or loads depending on weather conditions.
+We modeled the grid as a coupled oscillators network and approximated the dynamics with a system of second order linear differencial equations.
+We seek to control this dynamic when small perturbations occur by placing storage devices at a subset of the nodes.
+We believe that the energy that should be injected or absorbed from the grid by these controllers should be minimized.
+Given the diversity of possible situations, we seek the subset of nodes that, on average, necessitates minimum control energy.
+Based on the gramian matrix of the system and submodular set functions, we show how to find driver sets that use low energy on average and respect the system constraints.
+
+We believe that interesting work could be done by combining this model with real production and consumption data.
+There are indeed complex spatial and temporal correlations that impacts these distributions.
+Weather perturbations would therefore affect the nodes not completely at random, and affect how control should be designed.
+
+On the other hand, we seek in this paper, drivers that achieve full control of the system although this might not be needed in practice.
+More subtle controllability schemes might give better results both in terms of number of drivers and control energy.
 

### 2016-03-04 Article.tex
update

@@
 On the other hand, if the system is controllable, there might be more than one sequence of control inputs $u(t)$ that could drive it from $Y_0$ to $Y_f$.
-Among all these possibilities, optimal control is devoted to find the sequence that minimize some cost function J that can depend on the sates of the system $Y(t)$, the control inputs $u(t)$, and the final state $Y(T)$.
+Among all these possibilities, optimal control is devoted to find the sequence that minimize some cost function that can depend on the sates of the system $Y(t)$, the control inputs $u(t)$, and the final state $Y(T)$.
 As explained above, we are interested in the sequence that requires the minimum amount of control energy \cite{Yan2012}.
@@
 where $ \nu(T) = Y_f - A^T Y_0 $ is the difference between the desired final state and the final free state, and $W(T) = \sum_{k=0}^{T} A^k B B^T (A^T)^k $ is called the Gramian matrix of the system.
-The Gramian is a popular tool in control theory since it is deeply linked to the controllability of the system.
-It can indeed be shown that the system is controllable if and only if the Gramian is not singular, and that its rank indicates the dimension of the controllable subspace.
+It can be shown that the system is controllable if and only if the Gramian is not singular, and that its rank indicates the dimension of the controllable subspace.
 Besides, the minimum control energy associated with the inputs $ u^{\star}(t)$ can be written as :
@@
 \begin{equation}
+\label{eq:energy}
  \mathcal{E}_{min} = \nu(T)^T W^{-1}(T) \nu(T)
@@
 
-Not surprisingly, the control energy depends on the initial and final states as well as on the inverse of the Gramian.
-In our prosumer case, we want to find the best battery placement, whose quality should not depend on the initial and final states.
-Indeed, these states as well as  the power  distribution $ \Psi $ are susceptible to change, and we would like to perform well, on average, whatever the states and distributions.
-Therefore, we concentrate on  the Gramian as a way to quantify the average controllability.
+In cases where W is not invertible, the pseudo-inverse $W^{\dagger}$ can be used to obtain similar information in the controllable subpsace.
+The last row of matrix A in equation \ref{eq:final_with_control} adds a dimension to the state space that does not correspond to any phase angle or frequency but makes W singular.
+Since we seek control in the first $2N$ dimensions, we will use $W^{\dagger}$ in practice, and consider full control of the system if $rank[W]=2N$.
 
-Since the control energy is related to $ W^{-1} $, small energies are obtained from "large" W.
-In \cite{Summers2014}, the authors introduce several Gramian-based metrics that have concrete meanings in terms of the system controllability.
-For example, the trace of the inverse Gramian ($ Tr[ W^{-1}(T)] $) quantifies the average control energy necessary to move the system around the state space.
-These metrics measures some notions of the control allowed by a given driver node set.
-But they can be understood the other way around : what drivers should one select as to optimize a given control-based metric ?
+Not surprisingly, $\mathcal{E}_{min}$ depends on the initial and final states as well as on the inverse of W which only depends on A and B.
+The gramian can thus be used to obtain information about average controllability in the state space.
+In the prosumer network that we consider, generators and loads ($\Psi$) are suceptible to change, meaning that initial ($Y_0$) and final ($Y_f$) states might also vary.
+In such a scenario, we prefer to aim at good performances on average, rather than very good performances in few specific cases and bad performances in all other situations.
 
-Using the Gramian-based metric $ \mathcal{M} $, let $ F(S) =\mathcal{M}(W_S(T))$, where $ W_S(T) $  is the Gramian matrix based on the driver node set S.
-\cite{Summers2014} showed that for some metrics, F is a submodular set function.
-This means that, for these metrics, we can use the greedy heuristic discussed above in order to maximize F.
+Since the control energy is related to $ W^{-1} $, systems with "large" W will tend to be controlled with low energy.
+Still, this notion of "large" W is not very accurate and we need to quantify it by taking some metric based on W.
+We already mentioned the rank of W as being the dimension of the controllable subspace, but it is also known that the trace of W and $W^{-1}$ give information about average controllability and average control energy \cite{Summers2014}.
+In such a situation, we can build a set function that, given a set of driver nodes S, returns the value of one of these metrics.
+Indeed, given A and S, we can obtain the system dynamics $(A,B_S)$ and compute the gramian $W_S$.
+In other word, such a function would quantify the ability of a given set of nodes to contol the system on average.
+
+A key point, demonstrated in \cite{Summers2014}, is that the set function $ F:S \longrightarrow Tr[W_S] $ is modular and the two functions $F:S \longrightarrow Tr[W^{-1}_S] $ ans $ F:S \longrightarrow rank[W_S] $ are submodular.
+This nice result enables us to look for the driver node set that optimizes some notion of the average controllability using a simple greedy heuristic with a worst case guarantee (see section X).
+
+Theoretically, we are now able, for a given prosumer network, to compute A, B, and W.
+For some trajectory in the state space, we can use equation \ref{eq:ustar} to find the optimum control inputs.
+However, since we are considering a power grid as a network, we would like to verify that these inputs will ensure that the system physical constraints are respected in the process.
+We propose in the next section to take into account constraints related to line capacities and storage imperfections.
 
@@
 
-\subsection{Controllability Constraints}
-
- Considering the dynamic of equation \ref{eq:final_with_control}, we notice that the Gramian based on the transition matrix A is never invertible, meaning that the system should not be controllable.
+% Considering the dynamic of equation \ref{eq:final_with_control}, we notice that the Gramian based on the transition matrix A is never invertible, meaning that the system should not be controllable.
 It is easy to see that, indeed, we do not have full control because of the last row of the matrix.
@@
 
-Recall that the simplified dynamic (see section II) implies that, the power transmitted on line $ (i,j)$ is $ P_{i \longrightarrow j} = -P_{ij}^{MAX} \left( \theta_j(t) - \theta_i(t) \right) $.
-Therefore, we write the flow constraints at time t as :
+The power transmitted on line $ (i,j)$ is $ P_{i \longrightarrow j} = -P_{ij}^{MAX} \left( \theta_j(t) - \theta_i(t) \right) $. Flow constraints at time t can thus be written as :
 
 \begin{equation}
+\label{eq:flow_cons}
 \forall (i,j) \in V^2,\ \ g_{ij} \left|\ \theta_j(t) - \theta_i(t)\ \right| \leq 1
 \end{equation}
-If this constraint is verified for all instant t in the control time range, then the control inputs satisfy the flow constraints.
+If these constraints are verified for all instants t during the control phase then no line gets overloaded by the action of the control inputs.
 
-\subsection{Power Distribution Constraints}
-As explained above, we wish to obtain results that do not depend on the power distribution $ \Psi $ of the prosumers.
+%\subsection{Power Distribution Constraints}
+%As explained above, we wish to obtain results that do not depend on the power distribution $ \Psi $ of the prosumers.
 However, in order to achieve synchronization along with feasible power flows, we need to define some constraints on this distribution.
-As shown in [], a condition for achieving synchronisation in a coupled oscillators network is $ \Vert L^{\dagger}\omega \Vert_{\infty} \leq sin(\gamma) $, where L is the Laplacian matrix of the network, $ \omega $ is the vector of the natural frequencies of the oscillators, and $ \Vert x \Vert_{\infty} = max \left| x_i - x_j \right| $.
-If this condition is satisfied, the oscillators will synchronize at the common frequency $ \omega_{SYNC} $ with the phase lock $ g_{ij}\left| \theta_i - \theta_j \right| \leq    \gamma \in[0,\frac{\pi}{2}],\ \forall (i,j) \in V^2 $.
 
-Therefore, we can write the following synchronization constraint :
+\cite{Dorfler2013} showed that a condition for synchronisation in a coupled oscillators network is $ \Vert L^{\dagger}\omega \Vert_{\infty} \leq sin(\gamma) $, where L is the Laplacian matrix of the network, $ \omega $ is the vector of the natural frequencies of the oscillators, and $ \Vert x \Vert_{\infty} = max \left| x_i - x_j \right| $.
+If this condition is satisfied, the oscillators synchronize at the common frequency $ \omega_{SYNC} $ with the phase lock $ g_{ij}\left| \theta_i - \theta_j \right| \leq    \gamma \in[0,\frac{\pi}{2}],\ \forall (i,j) \in V^2 $.
+The synchronisation condition becomes in our settings :
 
@@
 Where $ A \circ B $ represents the Hadamard product between matrices A and B.
-If constraint \ref{synchro_constraint} is satisfied, the oscillators will synchronize to a common frequency $ \omega_{SYNC} = \frac{\sum_{k=1}^{N}\Psi_k}{\sum_{k=1}^{N}\alpha_k} = \frac{\sum_{k=1}^{N}P_{S,k}}{I\Omega N \alpha}$, and the flows at steady state will be feasible.
-Note that $\omega_{SYNC} $ should be zero in order to synchronize to the main frequency $\Omega$ (recall that the phase angles and frequencies represent deviations from $ \Omega$).
-This gives the following intuitive zero sum constraint :
+If constraint \ref{synchro_constraint} is satisfied, the oscillators synchronize to a common frequency $ \omega_{SYNC} = \frac{\sum_{k=1}^{N}\Psi_k}{\sum_{k=1}^{N}\alpha_k} = \frac{\sum_{k=1}^{N}P_{S,k}}{I\Omega N \alpha}$.
+
+Since the dynamics is expressed in term of deviations from $ \Omega$, synchronization at $ \Omega $ is achieved if $\omega_{SYNC} = 0 $. Wich gives the production consumption balance constraint :
 
@@
 
-Constraint \ref{zero_sum_constraint} simply states that production should match demand in order for the system to remain stable.
+Constraints \ref{synchro_constraint} and \ref{zero_sum_constraint} ensure that the system is able to synchronize at $\Omega$ without overloading lines.
 
@@
 
-Let i be a battery.
-In the simple model we adopt here, it is characterised by its maximum charge/discharge rate $r_i$, the amount of energy stored at time t $ \Lambda_i(t) $, and its maximum capacity $ \Lambda_{i, MAX} $.
-For simplicity, we will consider all maximum charge/discharge rates to be the same ($r$), and all maximum capacities to be the same $\Lambda_{Max}$.
-Furthermore, we denote by $ \Lambda(t) $ the vector of energy level at time t.
-
-Since the control input vector $ u^{\star}(t) $ specifies the control inputs, the energy level dynamic of the batteries can be written as :
-
-\begin{equation}
-\Lambda(t+\Delta t) = \Lambda(t) - u^{\star}(t) I \Omega 
-\end{equation}
-
+As for power lines, we need to take into account the imperfections of the storage that we consider.
+In practice, we could use some detailed and complex models, but as to keep things simple, we adopt here a straightforward approach.
+A storage i has a maximum charge/discharge rate $r_i$, some amount of energy stored at time t $ \Lambda_i(t) $, and a maximum capacity $ \Lambda_{i, MAX} $.
+We assume that all maximum rates ($r$) and all maximum capacities $\Lambda_{Max}$ are the same accross the storage equipments.
+We also denote by $ \Lambda(t) $ the vector of energy level at time t.
+Since $ u^{\star}(t) $ specifies the control inputs, the energy level dynamic of the batteries is $ \Lambda(t+\Delta t) = \Lambda(t) - u^{\star}(t) I \Omega $.
 Which can also be written as :
@@
 Where $ \Lambda(0) $ is the vector of initial levels in the batteries.
-
-Obviously, the energy level in a battery is bounded. This leads to the following set of constraints :
-
-\begin{equation}
- \forall t, \ \ 0 \leq \Lambda(t) \leq \Lambda_{MAX}
-\end{equation}
-
+Obviously, $\Lambda(t)$ has to stay within possible bounds : $ \forall t, \ \ 0 \leq \Lambda(t) \leq \Lambda_{MAX} $.
 Which can be written in terms of the control inputs :
@@
  \end{equation}
-
-If, for a given $ u^{\star} $, the constraints of eq. \ref{level_constraints} and \ref{rate_constraints} are satisfied, then the control inputs satisfy the storage constraints.
 

### 2016-03-07 Article.tex
update

@@
 \begin{equation}
- \ddot{ \theta_i } \sim \psi_i - \alpha \dot{ \theta_i } + \sum_{j\neq i} K_{ij} g_{ji} sin \left[ \theta_j - \theta_i \right] 
+ \ddot{ \theta_i } \sim \psi_i - \alpha \dot{ \theta_i } - \sum_{j\neq i} K_{ij} sin \left[ \theta_j - \theta_i \right] 
 \end{equation}
 
-Where $ \alpha = \frac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $, and $ g_{ij} $ is the coefficient of the adjacency matrix G.
+Where $ \alpha = \frac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $.
 By working in a rotating frame, we can simplify $ \psi_i=\frac{P_{S,i}}{I \Omega} $.
@@
 \label{eq:with_simplification}
-\ddot{\theta} = \Psi - \alpha \dot{\theta} - KL\theta
+\ddot{\theta} = \Psi - \alpha \dot{\theta} - (K \circ L)\theta
 \end{equation}
 
-Where L is the Laplacian matrix of the underlying topology ($ k_i $ is the degree of node i):
+Where $ A \circ B $ represents the Hadamard product between matrices A and B, and L is the Laplacian matrix of the underlying topology ($ k_i $ is the degree of node i):
 
@@
 \begin{equation}
- \dot{X} = \left( \begin{array}{cc} 0 & I \\ - KL & -\alpha I \end{array} \right) X + \left( \begin{array}{c} 0 \\ \Psi \end{array} \right)
+ \dot{X} = \left( \begin{array}{cc} 0 & I \\ - (K \circ L) & -\alpha I \end{array} \right) X + \left( \begin{array}{c} 0 \\ \Psi \end{array} \right)
 \end{equation}
@@
 
-With $ M = \left( \begin{array}{cc} I & I \Delta t \\ -KL \Delta t & (1-\alpha \Delta t) I \end{array} \right) $.
+With $ M = \left( \begin{array}{cc} I & I \Delta t \\ -(K \circ L) \Delta t & (1-\alpha \Delta t) I \end{array} \right) $.
 
@@
 \begin{equation}
- A =   \left( \begin{array}{ccc} I & I \Delta t & 0 \\ -KL \Delta t & (1-\alpha \Delta t)I & \Psi \Delta t \\ 0&0&1 \end{array} \right)
+ A =   \left( \begin{array}{ccc} I & I \Delta t & 0 \\ -(K \circ L) \Delta t & (1-\alpha \Delta t)I & \Psi \Delta t \\ 0&0&1 \end{array} \right)
 \end{equation}
@@
 Where the control matrix B is an indicator of the driver nodes, and $ u(t) $ represents the control vector input at time t.
+In other words, B maps the locations of the storage equipments in the system, and $u(t)$ can be seen as the power that needs to be injected or absorbed from the grid.
 The system is said to be controllable in T steps if it can be steered from any initial state $ Y_0 $ to any final state $ Y_f $ through a sequence of control inputs $ u(t) $.
 
-Controllability in complex network became popular with the well-known structural controllability framework introduced by Lin \cite{1100557}.
+Controllability in complex network became popular with the structural controllability framework introduced by Lin \cite{1100557}.
 This approach relates the search of the minimum driver set to a maximum matching in a bipartite graph, which can be done with the Hopcroft-Karp algorithm in $ O(|E|\sqrt{|V|}) $ in the worst case.
@@
 Note however that the independance of the parameters is a key assumption.
-If A is the adjacency matrix of an undirected network for instance, independance is not true since A is symetric, and one must use different tools to study controllability \cite{Yuan2014}.
+If A is the adjacency matrix of an undirected network for instance, independance is not verified (since A is symetric).
+In this case, one should use different tools to study controllability \cite{Yuan2014}.
 
 On the other hand, if the system is controllable, there might be more than one sequence of control inputs $u(t)$ that could drive it from $Y_0$ to $Y_f$.
-Among all these possibilities, optimal control is devoted to find the sequence that minimize some cost function that can depend on the sates of the system $Y(t)$, the control inputs $u(t)$, and the final state $Y(T)$.
+Among all these possibilities, optimal control is devoted to find the sequence that minimizes some cost function that can depend on the sates of the system $Y(t)$, the control inputs $u(t)$, and the final state $Y(T)$.
 As explained above, we are interested in the sequence that requires the minimum amount of control energy \cite{Yan2012}.
 Let $ \mathcal{E} = \int_{t_0}^{t_0+T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system.
-It can be shown \cite{Liu2015} that the control input sequence that minimizes the control energy can be written as :
+It can be shown \cite{Liu2015} that the control input sequence that minimizes $ \mathcal{E} $ can be written as :
 
@@
 where $ \nu(T) = Y_f - A^T Y_0 $ is the difference between the desired final state and the final free state, and $W(T) = \sum_{k=0}^{T} A^k B B^T (A^T)^k $ is called the Gramian matrix of the system.
-It can be shown that the system is controllable if and only if the Gramian is not singular, and that its rank indicates the dimension of the controllable subspace.
+It can also be shown that the system is controllable if and only if the Gramian is not singular, and that its rank indicates the dimension of the controllable subspace.
 Besides, the minimum control energy associated with the inputs $ u^{\star}(t)$ can be written as :
@@
 
-The power transmitted on line $ (i,j)$ is $ P_{i \longrightarrow j} = -P_{ij}^{MAX} \left( \theta_j(t) - \theta_i(t) \right) $. Flow constraints at time t can thus be written as :
+The power that flows on line $ (i,j) \in E$ can be written as $ P_{i \longrightarrow j} = -P_{ij}^{MAX} \left( \theta_j(t) - \theta_i(t) \right) $. Flow constraints at time t are straightforward :
 
@@
 \label{eq:flow_cons}
-\forall (i,j) \in V^2,\ \ g_{ij} \left|\ \theta_j(t) - \theta_i(t)\ \right| \leq 1
+\forall (i,j) \in E,\ \left|\ \theta_j(t) - \theta_i(t)\ \right| \leq 1
 \end{equation}
@@
 
-\cite{Dorfler2013} showed that a condition for synchronisation in a coupled oscillators network is $ \Vert L^{\dagger}\omega \Vert_{\infty} \leq sin(\gamma) $, where L is the Laplacian matrix of the network, $ \omega $ is the vector of the natural frequencies of the oscillators, and $ \Vert x \Vert_{\infty} = max \left| x_i - x_j \right| $.
-If this condition is satisfied, the oscillators synchronize at the common frequency $ \omega_{SYNC} $ with the phase lock $ g_{ij}\left| \theta_i - \theta_j \right| \leq    \gamma \in[0,\frac{\pi}{2}],\ \forall (i,j) \in V^2 $.
+\cite{Dorfler2013} showed that a condition for synchronisation in a coupled oscillators network is $ \Vert L^{\dagger}\omega \Vert_{\infty,E} \leq sin(\gamma) $, where L is the Laplacian matrix of the network, $ \omega $ is the vector of the natural frequencies of the oscillators, and $ \Vert x \Vert_{\infty,E} = max_{(i,j) \in E} \left| x_i - x_j \right| $.
+If this condition is satisfied, the oscillators synchronize at the common frequency $ \omega_{SYNC} $ with the phase lock $ \left| \theta_i - \theta_j \right| \leq    \gamma \in[0,\frac{\pi}{2}],\ \forall (i,j) \in E $.
 The synchronisation condition becomes in our settings :
@@
 \label{synchro_constraint}
-\Vert \left( L \circ P^{MAX} \right)^{\dagger} \Psi \Vert_{\infty} \leq sin(1)
+\Vert \left( L \circ K \right)^{\dagger} \Psi \Vert_{\infty} \leq sin(1)
 \end{equation}
 
-Where $ A \circ B $ represents the Hadamard product between matrices A and B.
 If constraint \ref{synchro_constraint} is satisfied, the oscillators synchronize to a common frequency $ \omega_{SYNC} = \frac{\sum_{k=1}^{N}\Psi_k}{\sum_{k=1}^{N}\alpha_k} = \frac{\sum_{k=1}^{N}P_{S,k}}{I\Omega N \alpha}$.
@@
 
-\subsection{Algorithm}
+\subsection{Optimization with grid constraints}
 
-We propose the following greedy algorithm in order to find the smallest set of locations that minimizes the average control energy and satisfy the constraints.
+\begin{figure}
+\includegraphics[scale=.38]{./figure_2/figure_2.pdf}
+\caption{At time $ t_k=4500 $, $P_i(t_k) = P_i(t_{k-1}) + \delta P$ for some random node i (perturbation).
+Consequently, the frequencies $ \dot{ \theta }_i $ deviate from the synchronized state.
+At time $ t_k + d = 5000$, optimal control inputs (see equation \ref{eq:ustar}) are injected such that the system is brought to the synchronized state at time $ t_k + d + T = 5500 $ (T is the control time).}
+
+%At this time, if the mismatch between production and consumption is still active, we maintain synchrony by using optimal control with $T=1$. }
+
+%Time series of the N frequencies $\omega_i$ (phase angles are not shown for clarity).
+$d$ is the delay between the perturbation and the begining of the control phase that bring the system bak in the synchronized state in T time steps.
+Here, $d$ was chosen quite large for the clarity of the figure.
+Note that, if the perturbation is still active at this point, the system still needs control in order to stay synchronized (batteries should still compensate the power imbalance)}
+\label{fig:frequecies}
+\end{figure}
+
+Finding the driver set of size k $S_k$ using the gramian $W_{S_k}$ of the system $(A,B_{S_k})$ does not require initial and final states.
+Besides, $W_{S_k}$ does not depend on the modified power distribution $ \Psi $ of the nodes.
+We are indeed looking for k controllers that perform well on average over all possible situations.
+On the contrary, the constraints derived above (eq.
+\ref{eq:flow_cons}, \ref{synchro_constraint}, \ref{level_constraints} and \ref{rate_constraints}) are bound to a particular trajectory in the state space.
+We indeed check wether the power grid can sustain the controlled dynamics when transitioning from a given initial state to a target final state.
+As we do not know what these states could be, we could use multiple scenarios and test wether $S_k$ can control the system without violation of constraints.
+In this paper, we use a slightly different approach.
+We consider that the system is initially at equilibrium (all elements are synchronized at $\Omega$) and that control will be necessary if a perturbation (power inbalance) occurs and takes the system out of equilibrium (see figure \ref{fig:frequecies}).
+Depending on how much time we need to start the control phase, the state of the system (the initial state for control) might be "somewhere around" the synchronized state.
+In order to test wether a set $S_k$ can control the system without violation of constraints, we sample initial states within some hypersphere centered on the synchronized state and check all the constraints.
+If for all trajectories, the constraints are respected, then we consider that $S_k$ enables the control of the system.
+
+Increasing k means that we deploy more strorage which increases the costs but tends to lower the energy required as we will see in the next section.
+Using the submodularity of the set functions introduced above, we can build a sequence of growing sets $S_1 \subset S_2 \subset ...
+\subset S_k $ and stop as soon as $S_k$ enables the control of the system for some k.
+
+%We propose the following greedy algorithm in order to find the smallest set of locations that minimizes the average control energy and satisfy the constraints.
 We assume here that the system is in a given initial state $ Y_0 $ and we want to reach a desired final state $ Y_f$  in $ T $ time steps.
@@
 	\begin{algorithmic}
-                \While{  \small $ Not\ \left\{ \begin{array}{lll} 
-                                              rank \left[ W_{S_k}(T) \right] < 2N\ and \\ 
-                                              \forall t,\forall (i,j) \in V^2, g_{ij} \left| \theta_j(t) - \theta_i(t) \right| \leq 1\  and \\
-                                              \forall t,-\frac{\Lambda(0)}{I \Omega} \leq -\sum_{k=0}^{t}u^{\star}(k) \leq \frac{ \Lambda_{MAX} - \Lambda(0)}{I \Omega}and \\ 
-                                              \forall t,\ \left| u^{\star}(t) \right| \leq \frac{r}{I\Omega}
-
-                                              \end{array} 
-                                              \right. $ }
-    \State $k\ \ \ \ \ \ \ \ \ \ \ \gets k+1$
-    \State $ S_k \ \ \ \ \ \ \ \ \ \gets Subopt(F, k)$
-    \State $ u^{\star}(t)\ \ \ \ \ \ \gets OptControl( S_{k} )$ 
+				\State $k=0$
+				\State $S_k = \emptyset $
+                \While{ Not Constrained control}
+                	\State $k \gets k+1$
+                	\State $ S_k = S_{k-1} \cup argmax_{i \in N \setminus S_{k-1} } F(S_{k-1} \cup \{i\}) $
                 \EndWhile
      \end{algorithmic}
-        \caption{Greedy submodular optimization with constraints}
+     \caption{Optimization with grid constraints}
         \label{algo_1}
 \end{algorithm}
+
+%\begin{algorithm}
+%        \begin{algorithmic}
+%                \While{  \small $ Not\ \left\{ \begin{array}{lll} 
+%                                              rank \left[ W_{S_k}(T) \right] < 2N\ and \\ 
+%                                              \forall t,\forall (i,j) \in V^2, g_{ij} \left| \theta_j(t) - \theta_i(t) \right| \leq 1\  and \\
+%                                              \forall t,-\frac{\Lambda(0)}{I \Omega} \leq -\sum_{k=0}^{t}u^{\star}(k) \leq \frac{ \Lambda_{MAX} - \Lambda(0)}{I \Omega}and \\ 
+%                                              \forall t,\ \left| u^{\star}(t) \right| \leq \frac{r}{I\Omega}
+%                                              
+%                                              \end{array} 
+%                                              \right. $ }
+%    \State $k\ \ \ \ \ \ \ \ \ \ \ \gets k+1$
+%    \State $ S_k \ \ \ \ \ \ \ \ \ \gets Subopt(F, k)$
+%    \State $ u^{\star}(t)\ \ \ \ \ \ \gets OptControl( S_{k} )$ 
+%\EndWhile
+%        \end{algorithmic}
+%        \caption{Greedy submodular optimization with constraints}
+%        \label{algo_1}
+%\end{algorithm}
 \normalsize
@@
 
-We first consider a small example of 10 nodes connected with an erdos-renyi topology.
+%We first consider a small example of 10 nodes connected with an erdos-renyi topology.
 The power distribution and the capacities of the lines are selected randomly such that the system is at equilibrium.
@@
 
-\begin{figure}
-\includegraphics[scale=.38]{./figure_2/figure_2.pdf}
-\caption{Time series of the N frequencies $\omega_i$ (phase angles are not shown for clarity).
-$d$ is the delay between the perturbation and the begining of the control phase that bring the system bak in the synchronized state in T time steps.
-Here, $d$ was chosen quite large for the clarity of the figure.
-Note that, if the perturbation is still active at this point, the system still needs control in order to stay synchronized (batteries should still compensate the power imbalance)}
-\label{fig:frequecies}
-\end{figure}
-
 The purpose of selecting the drivers according to a Gramian related metric is to minimize, on average, the amount of control energy needed.
 In other words, if we select the drivers randomly, we expect to need, on average, more energy to control the system.
-In figure \ref{fig:densities}, we draw $10^4$ example systems of 50 nodes with random scale-free topologies (Barabasi-Albert model is used), and random power distributions and line capacities.
+In figure \ref{fig:densities}, we draw $10^4$ example systems of 50 nodes with random scale-free topologies (Barabasi-Albert model), and random power distributions and line capacities.
 For each system we select a random number of drivers $N_D \sim \mathcal{U}(1,N_{nodes}) $ and we find two driver sets of size $N_D $.
 One is chosen randomly and the other is found with algorithm \ref{algo_1}.
-For both sets, we draw a random initial state $Y_i$ and a random final state $Y_f$, and we compute the control energy required for driving the system from $Y_i$ to $Y_f$.
+For both sets, if the control is possible, we draw a random initial state $Y_i$ and a random final state $Y_f$, and we compute the control energy required for driving the system from $Y_i$ to $Y_f$.
 On figure \ref{fig:densities} we plot each system as a dot with coordinates $(n_D = N_D / N_{nodes}, log(\mathcal{E}))$.
@@
 As visible, both curves decrease as the number of drivers increases, meaning that, as the number of drivers grows, the average control energy tends to decrease.
-Furthermore, we tend to use, as exected, less energy when the drivers are carefully selected.
+Furthermore, we tend to use, as expected, less energy when the drivers are carefully selected.
 Note that this difference tends to zero when $n_D$ tends to one, because almost all nodes are then selected, yielding little flexibility for optimization.
@@
 \includegraphics[scale=.35]{./figure_3/figure_3}
-\caption{Control energy $log(\mathcal{E})$ against size of control set $n_D$ for random and optimized driver sets.
+\caption{Logarithm of average control energy $log(\mathcal{E})$ against size of control set $n_D$ for random and optimized driver sets.
 Topologies are drawn from a Barabasi-Albert model with $ N_{nodes} = 50 $, power and line capacity distributions are random.
-For both case, $10^4$ instances are considered and the results are displayed as dots.
+$10^4$ instances are considered and the results are displayed as dots.
 The lines shows the average for both case.
@@
 \begin{itemize}
-\item Level 1 : full control : the system can be moved from any point to any other point of the state space. We do not consider constraints on the energy which can tend to infinity. 
-\item Level 2 : full control + line capacities constraints + battery constraints (see algorithm \ref{algo_1}).
+\item Level 1 : full control : the system can be moved from any point to any other point of the state space. We do not consider any other constraint. 
+\item Level 2 : full control + grid constraints (see section \ref{sec:control_and_submodularity}).
 This level is much more demanding since we impose the ability to move the system from any point to any other point, without overloading any line or breaking any battery limits in the process.
@@
 The top three curves show the results for the three metrics taken into consideration when all constraints are considered (see Level 2 in the main text).
-The bottom three curves exhibit the results for the same metrics when only the full controllability constraint is considered (see Level 1 in the main text).}
+The bottom three curves exhibit the results for the same metrics when only the full controllability constraint is considered (see Level 1).}
 \label{fig:erdos_renyi}
@@
 
-Because transmission power grids are spatially and politically constrained (each contry has its own transmission grid which is interconnected with neighboring grids at some specific locations), we study the control in clustered graphs.
+Because transmission power grids are spatially constrained (each contry has its own transmission grid which is interconnected with neighboring grids at some specific locations), we study the control in clustered graphs.
 More precisely, we use a block model $(N, N_{clusters}, p_{in}, p_{out} ) $ to generate random topologies where $ N $ is the number of nodes, $N_{clusters} $ is the number of clusters, $p_{in} $ is the probability that two nodes within the same cluster are linked, and $ p_{out} $ is the probability that two nodes in two distinct clusters are connected.

### 2016-03-17 Article.tex
new figure style

@@
 	\begin{center}
-	\includegraphics[scale=.3]{./figure_1/figure_1}
+	\includegraphics[scale=.25]{./figure_1/figure_1}
 	\caption{Submodularity example with sensor placements.}
@@
 \begin{figure}
-\includegraphics[scale=.38]{./figure_2/figure_2.pdf}
+\includegraphics[scale=.55]{plot5.pdf}%./figure_2/figure_2.pdf}
 \caption{At time $ t_k=4500 $, $P_i(t_k) = P_i(t_{k-1}) + \delta P$ for some random node i (perturbation).
@@
 \begin{figure}
-\includegraphics[scale=.35]{./figure_3/figure_3}
+\includegraphics[scale=.55]{plot4.pdf}%./figure_3/figure_3}
 \caption{Logarithm of average control energy $log(\mathcal{E})$ against size of control set $n_D$ for random and optimized driver sets.
@@
 \begin{figure}
-\includegraphics[scale=.25]{./figure_4/figure_4}
+\includegraphics[scale=.55]{plot3.pdf}%./figure_4/figure_4}
 \caption{ $ n_D $ against link probability p for erdos-renyi topologies ($ N=100 $).
@@
 
-\begin{figure}
-\includegraphics[scale=.25]{./figure_5/figure_5}
-\caption{$n_D$ against $N_{clusters} $ for different values of $ p_{in} $ and $ p_{out} = 0.1 $. $ N=200 $ and curves are averaged over 100 realizations.}
-\label{fig:block_model_1}
-\end{figure}
+\begin{figure*}
+	\begin{center}
+		\includegraphics[scale=.5]{Blockmodel.png}
+		\caption{left panel : $n_D$ against $N_{clusters} $ for different values of $ p_{in} $ and $ p_{out} = 0.1 $.
+$ N=200 $ and curves are averaged over 100 realizations.
+right panel : $n_D$ against $ p_{in} $ for systems of $N=200 $, $p_{out} = 0.1 $, and for different values of $N_{clusters} $.
+The curves are averaged over 100 realizations.}
+	\end{center}
+\end{figure*}
+
+%\begin{figure}
+%\includegraphics[scale=.25]{./figure_5/figure_5}
+%\caption{$n_D$ against $N_{clusters} $ for different values of $ p_{in} $ and $ p_{out} = 0.1 $. $ N=200 $ and curves are averaged over 100 realizations.}
+%\label{fig:block_model_1}
+%\end{figure}
 
@@
 
-\begin{figure}
-\includegraphics[scale=.25]{./figure_6/figure_6}
-\caption{$n_D$ against $ p_{in} $ for systems of $N=200 $, $p_{out} = 0.1 $, and for different values of $N_{clusters} $. The curves are averaged over 100 realizations.}
-\label{fig:block_model_2}
-\end{figure}
+%\begin{figure}
+%\includegraphics[scale=.25]{./figure_6/figure_6}
+%\caption{$n_D$ against $ p_{in} $ for systems of $N=200 $, $p_{out} = 0.1 $, and for different values of $N_{clusters} $. The curves are averaged over 100 realizations.}
+%\label{fig:block_model_2}
+%\end{figure}
 
@@
 \begin{figure}
-\includegraphics[scale=.37]{./figure_7/figure_7}
+\includegraphics[scale=.55]{plot2.pdf}%./figure_7/figure_7}
 \caption{$n_D$ against the relative standard deviation of the capacity distribution of the batteries $ \frac{ \sigma_{ \lambda } }{ \mu_{ \lambda} } $.

### 2016-03-17 figure_3/plot_figure_3.py

@@ -38,10 +38,11 @@ plt.scatter( 1./N_nodes * RAND[:,0], RAND[:,1], s= 60, color = '#FD6AF2', alpha 
 plt.scatter( 1./N_nodes * OPT[:,0], OPT[:,1], s= 60, color = '#FDAC86', alpha = .2 )
 plt.plot( 1./N_nodes * np.arange(1,N_nodes), MEANS, linewidth = 4, color = '#C703B8', label = 'random' )
 plt.plot( 1./N_nodes * np.arange(1,N_nodes), MEANS2, linewidth = 4, color = 'r', label = 'greedy optimization' )
-plt.xlim((0.4,1))
-plt.ylim((5,22))
+plt.xlim((0.45,1))
+plt.ylim((5,27))
 plt.xlabel(r'$n_D$', fontsize = 20)
 plt.ylabel(r'$log(\mathcal{E})$', fontsize = 20)
 plt.legend(loc=3, fontsize = 15)
 plt.grid()
 plt.savefig('./figure_3.pdf', dpi = 300)
+plt.savefig('./figure_3.png')
